### 2018-07-16 paper/paper.tex
Begin writing paper; add high level structure

%
% File naaclhlt2018.tex
%
%% Based on the style files for NAACL-HLT 2018, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
% \usepackage[showframe=true]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{lipsum}
\usepackage{url}
\usepackage{multirow}
\usepackage{changepage}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{amsmath}
\usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
% \hypersetup{draft}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}
\newcommand\confname{EMNLP 2018}
\newcommand\conforg{SIGDAT}
\newcommand{\colrule}{\noindent\rule{8cm}{0.4pt}}
\newcommand{\colruleend}{\noindent\rule{8cm}{2.0pt}}

% Avoid breaking inline equations ALWAYS (might break layout)
% https://tex.stackexchange.com/questions/14241/how-can-i-prevent-latex-from-breaking-inline-formulas-globally
\relpenalty=10000
\binoppenalty=10000

\title{IIIDYT at IEST 2018: Implicit Emotion Classification With Deep
Contextualized Word Representations}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\
  \And{}
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod
tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At
vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren,
no sea takimata sanctus est. Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut
labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata
sanctus est Lorem ipsum dolor sit amet.

\end{abstract}

\section{Introduction}
\begin{itemize}
    \item High level description of the task~\cite{klinger2018iest} and its \textit{raison d'être}
        (Francés para Edison)
    \item High level description of the model used and results obtained
\end{itemize}

\section{Related Work}
\begin{itemize}
    \item briefly describe previous Wassa papers~\cite{baziotis2018ntua,
        duppada2018seernet, abdou2018affecthor} (Should we do this despite not
        having based our work in any of them?)
    \item Talk about elmo~\cite{peters2018deep}
\end{itemize}

\section{Proposed Model}
\subsection{Preprocessing}
\begin{itemize}
    \item Replaced \texttt{\footnotesize[\#TRIGGERWORD\#]}, \texttt{\footnotesize@USERNAME},
        \texttt{\footnotesize[NEWLINE]}, and \texttt{\footnotesize http://url.removed}, by
        \texttt{\footnotesize\_\_TRIGGERWORD\_\_}, \texttt{\footnotesize\_\_USERNAME\_\_},
        \texttt{\footnotesize\_\_NEWLINE\_\_}, and \texttt{\footnotesize\_\_URL\_\_} respectively.
    \item Tokenized by using an emoji-aware modified version of the
        \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}}
        script. To incorporate emoji knowledge we used an emoji
        database\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}}
        which we also modified for avoiding conflict with emoji sharing unicode
        codes with common glyphs used in twitter such as the hashtag symbol.
        (Link to appendix for specifics; mention \texttt{keycaps})
\end{itemize}

\subsection{Model Architecture}

\begin{itemize}
    \item Elmo~\cite{peters2018deep} as word representations
    \item BiLSTM~\cite{graves2005framewise, graves2013speech} as context fine-tuner (?)
    \item max-pooling as word aggregation method
    \item slanted triangular learning rate schedule with mostly default params
        (just set the $\eta_{max}$ parameter to $0.001$.)
        as lr schedule~\cite{howard2018universal}
    \item 0.5 dropout after word layer, 0.1 after sentence layer~\cite{srivastava2014dropout}
    \item Only external feature is the elmo pre-trained language model
\end{itemize}

\subsection{Ensembling}

We trained several models with the previously-mentioned architecture, but
different initial parameters (by changing the random seed). We fed them the test
data, and averaged the predicted output probabilities. We tried all the possible
combinations of averaging $2$ to $10$ models, and found out that a specific
combination of $6$ models yielded the best results (show bar graph). This
provides evidence for the fact that having the same number of independent
classifiers as class labels provides the best results~\cite{bonab2016theoretical}.

\section{(Failed) experiments}
\begin{itemize}
    \item scalar gates, vector gates and elmo
    \item dropouts?
    \item POS tags
    \item combined pooling
    \item different learning rate schedules
    \item emoji vs no emoji
    \item sentence encoding lstm hidden size
    \item transformer architecture
    \item show different accuracies when using different training sizes?
\end{itemize}

\section{Conclusions and Future Work}

\bibliography{bibliography}
\bibliographystyle{acl_natbib_nourl}

\end{document}

### 2018-07-26 SentEval/senteval/sick.py
Add SentEval repo

@@ -1,0 +1,217 @@
+# Copyright (c) 2017-present, Facebook, Inc.
+# All rights reserved.
+#
+# This source code is licensed under the license found in the
+# LICENSE file in the root directory of this source tree.
+#
+
+'''
+SICK Relatedness and Entailment
+'''
+from __future__ import absolute_import, division, unicode_literals
+
+import os
+import io
+import logging
+import numpy as np
+
+from sklearn.metrics import mean_squared_error
+from scipy.stats import pearsonr, spearmanr
+
+from senteval.tools.relatedness import RelatednessPytorch
+from senteval.tools.validation import SplitClassifier
+
+
+class SICKRelatednessEval(object):
+    def __init__(self, task_path, seed=1111):
+        logging.debug('***** Transfer task : SICK-Relatedness*****\n\n')
+        self.seed = seed
+        train = self.loadFile(os.path.join(task_path, 'SICK_train.txt'))
+        dev = self.loadFile(os.path.join(task_path, 'SICK_trial.txt'))
+        test = self.loadFile(os.path.join(task_path, 'SICK_test_annotated.txt'))
+        self.sick_data = {'train': train, 'dev': dev, 'test': test}
+
+    def do_prepare(self, params, prepare):
+        samples = self.sick_data['train']['X_A'] + \
+                  self.sick_data['train']['X_B'] + \
+                  self.sick_data['dev']['X_A'] + \
+                  self.sick_data['dev']['X_B'] + \
+                  self.sick_data['test']['X_A'] + self.sick_data['test']['X_B']
+        return prepare(params, samples)
+
+    def loadFile(self, fpath):
+        skipFirstLine = True
+        sick_data = {'X_A': [], 'X_B': [], 'y': []}
+        with io.open(fpath, 'r', encoding='utf-8') as f:
+            for line in f:
+                if skipFirstLine:
+                    skipFirstLine = False
+                else:
+                    text = line.strip().split('\t')
+                    sick_data['X_A'].append(text[1].split())
+                    sick_data['X_B'].append(text[2].split())
+                    sick_data['y'].append(text[3])
+
+        sick_data['y'] = [float(s) for s in sick_data['y']]
+        return sick_data
+
+    def run(self, params, batcher):
+        sick_embed = {'train': {}, 'dev': {}, 'test': {}}
+        bsize = params.batch_size
+
+        for key in self.sick_data:
[…]

### 2018-07-26 SentEval/senteval/tools/relatedness.py

@@ -1,0 +1,134 @@
+# Copyright (c) 2017-present, Facebook, Inc.
+# All rights reserved.
+#
+# This source code is licensed under the license found in the
+# LICENSE file in the root directory of this source tree.
+#
+
+"""
+Semantic Relatedness (supervised) with Pytorch
+"""
+from __future__ import absolute_import, division, unicode_literals
+
+import copy
+import numpy as np
+
+import torch
+from torch import nn
+from torch.autograd import Variable
+import torch.optim as optim
+
+from scipy.stats import pearsonr
+
+
+class RelatednessPytorch(object):
+    # Can be used for SICK-Relatedness, and STS14
+    def __init__(self, train, valid, test, devscores, config):
+        # fix seed
+        np.random.seed(config['seed'])
+        torch.manual_seed(config['seed'])
+        assert torch.cuda.is_available(), 'torch.cuda required for Relatedness'
+        torch.cuda.manual_seed(config['seed'])
+
+        self.train = train
+        self.valid = valid
+        self.test = test
+        self.devscores = devscores
+
+        self.inputdim = train['X'].shape[1]
+        self.nclasses = config['nclasses']
+        self.seed = config['seed']
+        self.l2reg = 0.
+        self.batch_size = 64
+        self.maxepoch = 1000
+        self.early_stop = True
+
+        self.model = nn.Sequential(
+            nn.Linear(self.inputdim, self.nclasses),
+            nn.Softmax(),
+            )
+        self.loss_fn = nn.MSELoss()
+
+        if torch.cuda.is_available():
+            self.model = self.model.cuda()
+            self.loss_fn = self.loss_fn.cuda()
+
+        self.loss_fn.size_average = False
+        self.optimizer = optim.Adam(self.model.parameters(),
+                                    weight_decay=self.l2reg)
+
+    def prepare_data(self, trainX, trainy, devX, devy, testX, testy):
+        # Transform probs to log-probs for KL-divergence
+        trainX = torch.FloatTensor(trainX).cuda()
+        trainy = torch.FloatTensor(trainy).cuda()
+        devX = torch.FloatTensor(devX).cuda()
+        devy = torch.FloatTensor(devy).cuda()
+        testX = torch.FloatTensor(testX).cuda()
+        testY = torch.FloatTensor(testy).cuda()
+
+        return trainX, trainy, devX, devy, testX, testy
+
+    def run(self):
+        self.nepoch = 0
+        bestpr = -1
[…]

### 2018-07-26 SentEval/senteval/tools/validation.py
Fix "round" bug in SentEval validation

@@ -18,6 +18,8 @@ import logging
 import numpy as np
 from senteval.tools.classifier import MLP
 
+import torch
+
 import sklearn
 assert(sklearn.__version__ >= "0.18.0"), \
     "need to update sklearn to version >= 0.18.0"
@@ -100,7 +102,12 @@ class InnerKFoldClassifier(object):
                 clf = LogisticRegression(C=optreg, random_state=self.seed)
                 clf.fit(X_train, y_train)
 
-            self.testresults.append(round(100*clf.score(X_test, y_test), 2))
+            test_result = 100*clf.score(X_test, y_test)
+            if isinstance(test_result, torch.Tensor):
+                test_result = test_result.item()
+
+            test_result = round(test_result, 2)
+            self.testresults.append(test_result)
 
         devaccuracy = round(np.mean(self.devresults), 2)
         testaccuracy = round(np.mean(self.testresults), 2)

### 2018-07-26 SentEval/senteval/tools/validation.py
FIx other "round" errors in SentEval

@@ -183,6 +183,8 @@ class KFoldClassifier(object):
         yhat = clf.predict(self.test['X'])
 
         testaccuracy = clf.score(self.test['X'], self.test['y'])
+        if isinstance(testaccuracy, torch.Tensor):
+            testaccuracy = testaccuracy.item()
         testaccuracy = round(100*testaccuracy, 2)
 
         return devaccuracy, testaccuracy, yhat
@@ -226,8 +228,12 @@ class SplitClassifier(object):
             else:
                 clf = LogisticRegression(C=reg, random_state=self.seed)
                 clf.fit(self.X['train'], self.y['train'])
-            scores.append(round(100*clf.score(self.X['valid'],
-                                self.y['valid']), 2))
+
+            score = 100*clf.score(self.X['valid'], self.y['valid'])
+            if isinstance(score, torch.Tensor):
+                score = score.item()
+            scores.append(round(score, 2))
+
         logging.info([('reg:'+str(regs[idx]), scores[idx])
                       for idx in range(len(scores))])
         optreg = regs[np.argmax(scores)]
@@ -249,5 +255,7 @@ class SplitClassifier(object):
             clf.fit(self.X['train'], self.y['train'])
 
         testaccuracy = clf.score(self.X['test'], self.y['test'])
+        if isinstance(testaccuracy, torch.Tensor):
+            testaccuracy = testaccuracy.item()
         testaccuracy = round(100*testaccuracy, 2)
         return devaccuracy, testaccuracy

### 2018-07-30 paper/paper.tex
Update Paper, bibliografy, rsync_ignore
Update bibliography, write some intro paragraphs

@@
 \usepackage{amsmath}
+\usepackage{graphicx}
 \usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
@@
 
+% Style for acronyms; avoids splitting them.
+% See https://tex.stackexchange.com/questions/48435/how-do-i-prevent-tex-from-hyphenating-acronyms
+\newcommand{\Acronym}[1]{\mbox{#1}}
+
 \begin{document}
@@
 \section{Introduction}
-\begin{itemize}
-    \item High level description of the task~\cite{klinger2018iest} and its \textit{raison d'être}
-        (Francés para Edison)
-    \item High level description of the model used and results obtained
-\end{itemize}
+
+% FIXME: EDIT THIS SECTION TO MAKE PARAGRAPHS MORE COHESIVE
+Sentiment Analysis, also known as Opinion Mining, is a discipline whose
+objective is to automatically identify sentiment in written text
+\cite{balazs2016opinion}.
+
+The \Acronym{WASSA} 2018 Implicit Emotion Shared Task
+(\Acronym{IEST})~\cite{klinger2018iest}, consists in predicting the emotion of a
+word that has been excluded from a tweet. Its aim is to find ways to
+automatically learn the link between situations and the emotion they trigger.
+
+This problem can be reduced to sentence classification, in which we
+want to classify a sentence, or more specifically a tweet, into one of several
+categories.
+
+In this paper we present the system that obtained second place, proposed by the
+team \Acronym{IIIDYT} at \Acronym{IEST}\@. Our system was composed by a single
+pre-trained \Acronym{ELMo} layer for encoding words~\cite{peters2018deep}, a
+Bidirectional Long-Short Memory Network
+(\Acronym{BiLSTM})~\cite{graves2005framewise, graves2013speech}, for enriching
+word representations with context, a max-pooling operation for creating sentence
+representations from said word vectors, and finally a Dense Layer for projecting
+the sentence representations into label space (See
+Figure~\ref{fig:architecture}).
 
@@
 \begin{itemize}
-    \item briefly describe previous Wassa papers~\cite{baziotis2018ntua,
+    \item briefly describe previous Semeval papers~\cite{baziotis2018ntua,
         duppada2018seernet, abdou2018affecthor} (Should we do this despite not
         having based our work in any of them?)
+    \item mention previous Wassa papers
+    \item talk about some sentence classification tasks? Works that fall into
+        the broad category of mapping sentences to labels.
     \item Talk about elmo~\cite{peters2018deep}
@@
         \texttt{\footnotesize\_\_TRIGGERWORD\_\_}, \texttt{\footnotesize\_\_USERNAME\_\_},
-        \texttt{\footnotesize\_\_NEWLINE\_\_}, and \texttt{\footnotesize\_\_URL\_\_} respectively.
+        \texttt{\footnotesize\_\_NEWLINE\_\_}, and
+        \texttt{\footnotesize\_\_URL\_\_} respectively.
     \item Tokenized by using an emoji-aware modified version of the
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{images/best_ensembles.eps}
+\caption{Best Macro F1 vs. Number of Ensembled Models}
+\label{fig:}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{images/average_ensembles.eps}
+\caption{Best Macro F1 vs. Number of Ensembled Models}
+\label{fig:}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{images/iest_architecture.pdf}
+    \caption{Proposed architecture}
+    \label{fig:architecture}
+\end{figure}
+
 \section{(Failed) experiments}
@@
 
+\begin{table}[!h]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{lcc}
+
+        \textbf{Variation} & \textbf{Accuracy (\%)} & $\bf{\Delta}$\textbf{\%} \\
+        \hline
+        \hline
+        Submitted          & \textbf{69.23}         & -               \\
+        \hline
+        No emoji           & 68.36                  & - 0.87          \\
+        \hline
+        No ELMo            & 65.52                  & - 3.71          \\
+        \hline
+        Concat Pooling     & 68.47                  & - 0.76          \\
+        \hline
+        LSTM hidden=4096   & 69.10                  & - 0.13          \\
+        LSTM hidden=1024   & 68.93                  & - 0.30          \\
+        LSTM hidden=512    & 68.43                  & - 0.80          \\
+        \hline
+        POS emb dim=100    & 68.99                  & - 0.24          \\
+        POS emb dim=75     & 68.61                  & - 0.62          \\
+        POS emb dim=50     & 69.33                  & + 0.10          \\
+        POS emb dim=25     & 69.21                  & - 0.02          \\
+
+    \end{tabular}
+    \caption{Ablation study results}
+
+\end{table}
+
 \section{Conclusions and Future Work}

### 2018-07-31 get_data.sh
Get and preprocess true test data and labels

@@ -3,14 +3,17 @@
 USER=$1
 
 cd data
-wget --user=$USER --ask-password http://implicitemotions.wassa2018.com/data/protected/train-v3.csv.gz\
-                                 http://implicitemotions.wassa2018.com/data/protected/test.csv.gz
+wget --user=$USER --ask-password http://implicitemotions.wassa2018.com/data/protected/train-v3.csv.gz
 wget http://implicitemotions.wassa2018.com/data/unprotected/trial-v3.csv.gz
 wget http://implicitemotions.wassa2018.com/data/unprotected/trial-v3.labels.gz
 
+wget http://implicitemotions.wassa2018.com/data/unprotected/test-text-labels.csv.gz
+mv test-text-labels.csv.gz test.csv.gz
+
 gunzip train-v3.csv.gz
-gunzip test.csv.gz
 gunzip trial-v3.csv.gz
 gunzip trial-v3.labels.gz
+gunzip test.csv.gz
+gunzip test.labels.gz
 
 cd ..

### 2018-07-31 preprocess.sh

@@ -22,7 +22,7 @@ CLEAN_DEV_DATA_PATH="$PREPROCESSED_DATA_DIR/dev.csv"
 DEV_LABELS_PATH="$PREPROCESSED_DATA_DIR/dev_labels.csv"
 
 CLEAN_TEST_DATA_PATH="$PREPROCESSED_DATA_DIR/test.csv"
-TEST_LABELS_PATH="$PREPROCESSED_DATA_DIR/test_labels_fake.csv"
+TEST_LABELS_PATH="$PREPROCESSED_DATA_DIR/test_labels.csv"
 
 #######################################################################
 #                            No-emoji data                            #

### 2018-07-31 run.py
Save sentence representations when running test

@@ -98,6 +98,11 @@ arg_parser.add_argument("--warmup_iters", "-wup", default=4000, type=int,
 arg_parser.add_argument("--test", action="store_true",
                         help="Run this script in test mode")
 
+arg_parser.add_argument("--save_sent_reprs", "-ssr", action="store_true",
+                        default=False,
+                        help='Save sentence representations in the experiment '
+                        'directory. This is intended to be used with the --test flag')
+
 arg_parser.add_argument("--pos_emb_dim", "-pem", default=None, type=int,
                         help="The dimension to use for the POS embeddings. "
                              "If None, POS tags will not be used")
@@ -280,6 +285,16 @@ def main():
         with open(labels_filepath, 'w', encoding='utf-8') as f:
             f.writelines(labels)
         print(f'Saved prediction file in {labels_filepath}')
+
+        representations_filepath = os.path.join(
+            ext_experiment_path,
+            'sentence_representations.txt'
+        )
+
+        if hp.save_sent_reprs:
+            with open(representations_filepath, 'w', encoding='utf-8') as f:
+                np.savetxt(representations_filepath, eval_dict['sent_reprs'],
+                           delimiter=' ', fmt='%.8f')
         exit()
 
     writer = None

### 2018-07-31 src/config.py

@@ -34,7 +34,7 @@ DEV_LABELS = os.path.join(PREPROCESSED_DATA_PATH, 'dev_labels.csv')
 
 TEST = os.path.join(PREPROCESSED_DATA_PATH, 'test_no_emojis.csv')
 TEST_EMOJI = os.path.join(PREPROCESSED_DATA_PATH, 'test.csv')
-TEST_LABELS = os.path.join(PREPROCESSED_DATA_PATH, 'test_labels_fake.csv')
+TEST_LABELS = os.path.join(PREPROCESSED_DATA_PATH, 'test_labels.csv')
 
 TRAIN_POS = os.path.join(PREPROCESSED_DATA_PATH, 'train.csv.tagged.pos')
 DEV_POS = os.path.join(PREPROCESSED_DATA_PATH, 'dev.csv.tagged.pos')

### 2018-07-31 src/models/iest.py

@@ -563,6 +563,7 @@ class IESTClassifier(nn.Module):
 
         logits = self.dense_layer(sent_vec)
 
-        ret_dict = {'logits': logits}
+        ret_dict = {'logits': logits,
+                    'sent_reprs': sent_vec}
 
         return ret_dict

### 2018-07-31 src/train.py

@@ -81,15 +81,18 @@ class Trainer(object):
         num_batches = dev_batches.num_batches
         outputs = []
         true_labels = []
+        sent_reprs = []
         tqdm.write("Evaluating...")
         for batch_index in range(num_batches):
             batch = dev_batches[batch_index]
             out = self.model(batch)
-            # FIXME: Shouldn't call data()
+
             outputs.append(out['logits'].cpu().data.numpy())
+            sent_reprs.append(out['sent_reprs'].cpu().data.numpy())
             true_labels.extend(batch['labels'])
 
         output = np.vstack(outputs)
+        sent_reprs = np.vstack(sent_reprs)
         pred_labels = output.argmax(axis=1)
         true_labels = np.array(true_labels)
         tqdm.write(classification_report(true_labels, pred_labels,
@@ -106,5 +109,6 @@ class Trainer(object):
         pred_labels = [config.ID2LABEL[label] for label in pred_labels]
         ret_dict = {'accuracy': accuracy,
                     'labels': pred_labels,
-                    'output': output}
+                    'output': output,
+                    'sent_reprs': sent_reprs}
         return ret_dict

### 2018-08-27 paper/paper.tex
Update paper and bib

--- 2018-07-16 paper/paper.tex
@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{siunitx}
 \usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+\aclfinalcopy % Uncomment this line for the final submission
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
@@
 
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\
-  \And{}
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\}
+\author{Jorge A. Balazs, Edison Marrese-Taylor \and Yutaka Matsuo \\
+  Graduate School of Engineering\\
+  The University of Tokyo\\
+  {jorge, emarrese, matsuo@weblab.t.u-tokyo.ac.jp} \\
+}
 
 \date{}
+
+% Style for acronyms; avoids splitting them.
+% See https://tex.stackexchange.com/questions/48435/how-do-i-prevent-tex-from-hyphenating-acronyms
+\newcommand{\Acronym}[1]{\mbox{#1}}
 
@@
 \begin{abstract}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod
-tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At
-vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren,
-no sea takimata sanctus est. Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut
-labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata
-sanctus est Lorem ipsum dolor sit amet.
-
+In this paper we describe our system designed for the WASSA 2018 Implicit Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 26 teams with a test macro F1 score of $0.710$.
+The system is composed of a single pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory Network BiLSTM for enriching word representations with context, a max-pooling operation for creating sentence representations from said word vectors, and a Dense Layer for projecting the sentence representations into label space.
+Our official submission was obtained by ensembling 6 of these models initialized with different random seeds.
 \end{abstract}
@@
 \section{Introduction}
-\begin{itemize}
-    \item High level description of the task~\cite{klinger2018iest} and its \textit{raison d'être}
-        (Francés para Edison)
-    \item High level description of the model used and results obtained
-\end{itemize}
 
-\section{Related Work}
-\begin{itemize}
-    \item briefly describe previous Wassa papers~\cite{baziotis2018ntua,
-        duppada2018seernet, abdou2018affecthor} (Should we do this despite not
-        having based our work in any of them?)
-    \item Talk about elmo~\cite{peters2018deep}
-\end{itemize}
+Although the definition of emotion is still debated among the scientific community, the automatic identification and understanding of human emotions by machines has long been of interest in computer science.
+It has usually been assumed that emotions are triggered by the interpretation of a stimulus event according to its meaning.
 
-\section{Proposed Model}
+As language usually reflects the emotional state of an individual, it is natural to study human emotions by understanding how they are reflected in text.
+We see that many words indeed have affect as a core part of their meaning, for example, \textit{dejected} and \textit{wistful} denote some amount of sadness, and are thus associated with sadness.
+On the other hand, some words are associated with affect even though they do not denote affect.
+For example, \textit{failure} and \textit{death} describe concepts that are usually accompanied by sadness and thus they denote some amount of sadness.
+In this context, the task of automatically recognizing emotions from text has recently attracted the attention of researchers in Natural Language Processing.
+This task is usually formalized as the classification of words, phrases, or documents into predefined discrete emotion categories or dimensions.
+Some approaches have aimed at also predicting the degree to which an emotion is expressed in text \cite{wassa_emoint_2017}.
+
+In light of this, the \Acronym{WASSA} 2018 Implicit Emotion Shared Task (\Acronym{IEST}) \cite{klinger2018iest} was proposed to help find ways to automatically learn the link between situations and the emotion they trigger.
+The task consisted in predicting the emotion of a word excluded from a tweet.
+Removed words, or \textit{trigger-words}, included the terms ``sad'', ``happy'', ``disgusted'', ``surprised'', ``angry'', ``afraid'' and their synonyms, and the task was to predict the emotion they conveyed, specifically sadness, joy, disgust, surprise, anger and fear.
+
+From a machine learning perspective, this problem can be seen as sentence classification, in which the goal is to classify a sentence, or in particular a tweet, into one of several categories.
+In the case of \Acronym{IEST}, the problem is specially challenging since tweets contain informal language, the heavy usage of emoji, hashtags and username mentions.
+
+In this paper we describe our system designed for \Acronym{IEST}, which obtained the second place out of 26 teams.
+Our system did not require manual feature engineering and only minimal use of external data.
+Concretely, our approach is composed of a single pre-trained \Acronym{ELMo} layer for encoding words \cite{peters2018deep}, a Bidirectional Long-Short Memory Network (\Acronym{BiLSTM}) \cite{graves2005framewise, graves2013speech}, for enriching word representations with context, a max-pooling operation for creating sentence representations from said word vectors, and finally a Dense Layer for projecting the sentence representations into label space.
+To the best of our knowledge, our system, which we plan to release, is the first to utilize ELMo for emotion recognition.
+
+%  \item Code will be made available
+
+% FIXME: EDIT THIS SECTION TO MAKE PARAGRAPHS MORE COHESIVE
+% Sentiment Analysis, also known as Opinion Mining, is a discipline whose
+% objective is to automatically identify sentiment in written text
+% \cite{balazs2016opinion}.
+
+\begin{figure*}
+    \centering
+    \includegraphics[scale=0.6]{images/iest_architecture.pdf}
+    \caption{Proposed architecture.}
+    \label{fig:architecture}
+\end{figure*}
+
+\section{Proposed Approach}
+
 \subsection{Preprocessing}
-\begin{itemize}
-    \item Replaced \texttt{\footnotesize[\#TRIGGERWORD\#]}, \texttt{\footnotesize@USERNAME},
-        \texttt{\footnotesize[NEWLINE]}, and \texttt{\footnotesize http://url.removed}, by
-        \texttt{\footnotesize\_\_TRIGGERWORD\_\_}, \texttt{\footnotesize\_\_USERNAME\_\_},
-        \texttt{\footnotesize\_\_NEWLINE\_\_}, and \texttt{\footnotesize\_\_URL\_\_} respectively.
-    \item Tokenized by using an emoji-aware modified version of the
+
+As our model is purely character-based, we performed little data preprocessing. Table \ref{table:substitutions} shows the special tokens found in the datasets, and how we substituted them.
+
+\begin{table}[!h]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{ll}
+
+        \textbf{Original} & \textbf{Replacement} \\
+        \hline
+        \hline
+        \texttt{\footnotesize[\#TRIGGERWORD\#]} & \texttt{\footnotesize\_\_TRIGGERWORD\_\_} \\
+        \texttt{\footnotesize@USERNAME} & \texttt{\footnotesize\_\_USERNAME\_\_} \\
+        \texttt{\footnotesize[NEWLINE]} & \texttt{\footnotesize\_\_NEWLINE\_\_} \\
+        \texttt{\footnotesize http://url.removed} & \texttt{\footnotesize\_\_URL\_\_} \\
+
+    \end{tabular}
+    \caption{Preprocessing substitutions.}
+    \label{table:substitutions}
+\end{table}
+
+    Furthermore, we tokenized the text using a variation of the
     \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}}
-        script. To incorporate emoji knowledge we used an emoji
-        database\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}}
-        which we also modified for avoiding conflict with emoji sharing unicode
-        codes with common glyphs used in twitter such as the hashtag symbol.
-        (Link to appendix for specifics; mention \texttt{keycaps})
-\end{itemize}
+    script, a Python port of the original \texttt{Twokenize.java}
+    \cite{gimpel-EtAl:2011:ACL-HLT2011}. Concretely, we created an emoji-aware
+    version of it by incorporating knowledge from an emoji
+    database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}}
+    which we slightly modified for avoiding conflict with emoji sharing unicode
+    codes with common glyphs used in Twitter,\footnote{For example, the hashtag
+    emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3},
+which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and for making it compatible with Python 3.
+% \begin{itemize}
+%     \item briefly describe previous Semeval papers~\cite{baziotis2018ntua,
+%         duppada2018seernet, abdou2018affecthor} (Should we do this despite not
+%         having based our work in any of them?)
+%     \item mention previous Wassa papers
+%     \item talk about some sentence classification tasks? Works that fall into
+%         the broad category of mapping sentences to labels.
+%     \item Talk about elmo~\cite{peters2018deep}
+% \end{itemize}
 
-\subsection{Model Architecture}
+\subsection{Architecture}
 
-\begin{itemize}
-    \item Elmo~\cite{peters2018deep} as word representations
-    \item BiLSTM~\cite{graves2005framewise, graves2013speech} as context fine-tuner (?)
-    \item max-pooling as word aggregation method
-    \item slanted triangular learning rate schedule with mostly default params
-        (just set the $\eta_{max}$ parameter to $0.001$.)
-        as lr schedule~\cite{howard2018universal}
-    \item 0.5 dropout after word layer, 0.1 after sentence layer~\cite{srivastava2014dropout}
-    \item Only external feature is the elmo pre-trained language model
-\end{itemize}
+Figure \ref{fig:architecture} summarizes our proposed architecture.
+Our input is based on Embeddings from Language Models (ELMo) by \citet{peters2018deep}.
+These are character-based word representations allowing the model to avoid the ``unknown token'' problem.
+ELMo uses a set of convolutional neural networks to extract features from character embeddings, and builds word vectors from them.
+These are then fed to a multi-layer Bidirectional Language Model (BiLM) which returns context-sensitive vectors for each input word.
 
-\subsection{Ensembling}
+We used a single-layer BiLSTM as context fine-tuner \cite{graves2005framewise, graves2013speech}, on top of the ELMo embeddings, and then aggregated the hidden states it returned by using max-pooling, which has been shown to perform well on sentence classification tasks \cite{conneau2017supervised}.
 
-We trained several models with the previously-mentioned architecture, but
-different initial parameters (by changing the random seed). We fed them the test
-data, and averaged the predicted output probabilities. We tried all the possible
-combinations of averaging $2$ to $10$ models, and found out that a specific
-combination of $6$ models yielded the best results (show bar graph). This
-provides evidence for the fact that having the same number of independent
-classifiers as class labels provides the best results~\cite{bonab2016theoretical}.
+Finally, we used a single-layer fully-connected network for projecting the pooled BiLSTM output into a vector corresponding to the label logits for each predicted class.
 
-\section{(Failed) experiments}
-\begin{itemize}
-    \item scalar gates, vector gates and elmo
-    \item dropouts?
-    \item POS tags
-    \item combined pooling
-    \item different learning rate schedules
-    \item emoji vs no emoji
-    \item sentence encoding lstm hidden size
-    \item transformer architecture
-    \item show different accuracies when using different training sizes?
-\end{itemize}
+\subsection{Implementation Details and Hyperparameters}
+
+\hspace{\parindent} \textbf{ELMo Layer}: We used the official AllenNLP implementation of the ELMo model\footnote{\tiny\url{https://allenai.github.io/allennlp-docs/api/allennlp.modules.elmo.html}}, with the official weights pre-trained on the 1 Billion Word Language Model Benchmark, which contains about 800M tokens of news crawl data from WMT 2011 \cite{Chelba2014}.
+
+\textbf{Dimensionalities}: By default the ELMo layer outputs a 2048-dim vector, which we then feed to a BiLSTM resulting in a 4096-dim vector for each word of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read the input in opposite directions and whose outputs are concatenated at the sequence dimension.
+This results in a vector double the dimension of the input for each time step.}.
+After max-pooling the BiLSTM output over the sequence dimension we have a single 4096-dim vector corresponding to the tweet representation.
+This representation is finally fed to a single-layer fully-connected network with input size 4096, 512 hidden units, output size 6, and ReLU nonlinearities after the input and hidden layers.
+The output of the dense layer is a 6-dim logit vector for each example.
+
+\textbf{Loss Function}: As this corresponds to a multiclass classification problem (predicting a single class for each example, with more than 2 classes to choose from), we used the Cross-Entropy Loss as implemented in PyTorch \cite{paszke2017automatic}.
+
+\textbf{Optimization}: We optimized the model with Adam \cite{DBLP:journals/corr/KingmaB14}, using a slanted triangular learning rate schedule \cite{howard2018universal}, with default hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate $\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained by multiplying the number of epochs ($10$) times the total number of batches, which for the training dataset corresponds to $2396$ batches of $64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.
+
+\textbf{Regularization}: we used a dropout layer \cite{srivastava2014dropout}, with probability of 0.5 after both the ELMo and the hidden fully-connected layer, and another one with probability of 0.1 after the max-pooling aggregation layer.
+We also reshuffled the training examples between epochs, resulting in a different batch for each iteration.
+
+\textbf{Model Selection}: To choose the best hyperparameter configuration we measured the classification accuracy on the validation set (trial set).
+
+\subsection{Ensembles}
+
+Once we found the best-performing configuration we trained 10 models using different random seeds, and tried averaging the output probabilities of all their possible $\sum_{k=1}^{10}{\binom{10}{k}}=1023$ combinations.
+As Figure \ref{fig:best_ensembles} shows, we empirically found that a specific combination of $6$ models yielded the best results ($70.52\%$), providing evidence for the fact that using a number of independent classifiers equal to the number of class labels provides the best results when doing average ensembling \cite{bonab2016theoretical}.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=\columnwidth]{images/best_ensembles.eps}
+\caption{Best Macro F1 vs. Number of Ensembled Models.}
+\label{fig:best_ensembles}
+\end{figure}
+
+\section{Experiments and Analyses}
+
+We performed several experiments to gain insights on how the proposed model's performance interacts with the shared task's data.
+We performed an ablation study to see how some of the main hyperparameters affect performance, and an analysis of tweets containing hashtags and emoji to understand how these two types of tokens help the model predict the trigger-word's emotion.
+We also observed the effects of varying the amount of data used for training the model to evaluate whether it would be worthwhile to gather more training data.
+
+\subsection{Ablation Study}%
+\label{sub:ablation_study}
+
+We performed an ablation study on a single model having obtained 69.23\% accuracy on the validation set. Results are summarized in Table \ref{table:ablation}.
+
+\begin{table}[!h]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{lcc}
+
+        \textbf{Variation} & \textbf{Accuracy (\%)} & $\bf{\Delta}$\textbf{\%} \\
+        \hline
+        \hline
+        Submitted          & \textbf{69.23}         & -                        \\
+        \hline
+        No emoji           & 68.36                  & - 0.87                   \\
+        \hline
+        No ELMo            & 65.52                  & - 3.71                   \\
+        \hline
+        Concat Pooling     & 68.47                  & - 0.76                   \\
+        \hline
+        LSTM hidden=4096   & 69.10                  & - 0.13                   \\
+        LSTM hidden=1024   & 68.93                  & - 0.30                   \\
+        LSTM hidden=512    & 68.43                  & - 0.80                   \\
+        \hline
+        POS emb dim=100    & 68.99                  & - 0.24                   \\
+        POS emb dim=75     & 68.61                  & - 0.62                   \\
+        POS emb dim=50     & 69.33                  & + 0.10                   \\
+        POS emb dim=25     & 69.21                  & - 0.02                   \\
+        \hline
+        SGD optim lr=1     & 64.33                  & - 4.90                   \\
+        SGD optim lr=0.1   & 66.11                  & - 3.12                   \\
+        SGD optim lr=0.01  & 60.72                  & - 8.51                   \\
+        SGD optim lr=0.001 & 30.49                  & - 38.74                  \\
+
+    \end{tabular}
+    \caption{Ablation study results.}
+    \label{table:ablation}
+
+\end{table}
+
+Accuracies were obtained from the
+validation dataset.
+Each model was trained with the same random seed and hyperparameters, save for the one listed.
+``No emoji'' is the same model trained on the training dataset with no emoji, ``No ELMo'' corresponds to having switched the ELMo word encoding layer with a simple pre-trained GloVe embedding lookup, and ``Concat Pooling'' obtained sentence representations by using the pooling method described by \citet{howard2018universal}.
+``LSTM hidden'' corresponds to the the hidden dimension of the BiLSTM, ``POS emb dim'' to the dimension of the part-of-speech embeddings, and ``SGD optim lr'' to the learning rate used while optimizing with the schedule described by \citet{conneau2017supervised}.
+
+We can observe that the architectural choice that had the greatest impact on our model was the ELMo layer, providing a $3.71\%$ boost in performance as compared to using GloVe pre-trained word embeddings.
+
+We can further see that emoji also contributed significantly to the model's performance. In Section \ref{sub:effect_of_emoji_and_hashtags} we give some pointers to understanding why this is so.
+
+Additionally, we tried using the concatenation of the max-pooled, average-pooled and last hidden states of the BiLSTM as the sentence representation, following \citet{howard2018universal}, but found out that this impacted performance negatively.
+We hypothesize this is due to tweets being too short for needing such a rich representation.
+Also, the size of the concatenated vector was $4096\times3=12,288$, which probably could not be properly exploited by the 512-dimensional fully-connected layer.
+
+Using a greater BiLSTM hidden size did not help the model, probably because of the reason mentioned early; the fully-connected layer was not big or deep enough to exploit the additional information.
+Similarly, using a smaller hidden size neither helped.
+
+We found that using 50-dim part-of-speech embeddings slightly improved results, which implies that better fine-tuning this hyperparameter, or using a better POS tagger could yield an even better performance.
+
+Regarding optimization strategies, we also tried using SGD with different learning rates and a step-wise learning rate schedule as described by \citet{conneau2018}, but we found that doing this did not improve performance.
+
+Finally, Figure \ref{fig:dropouts} shows the effect of using different dropout probabilities.
+We can see that having higher dropout after the word-representation layer and the fully-connected network's hidden layer, while having a low dropout after the sentence encoding layer yielded better results overall.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=\columnwidth]{images/dropout_table.pdf}
+\caption{Dropout Variations.
+Rows correspond to the dropout applied both after the ELMo layer (word encoding layer) and after the fully-connected network's hidden layer, while columns correspond to the dropout applied after the max-pooling operation (sentence encoding layer.}
+\label{fig:dropouts}
+\end{figure}
+
+\subsection{Effect of Emoji and Hashtags}%
+\label{sub:effect_of_emoji_and_hashtags}
+
+\begin{table}[!h]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{lcc}
+
+        \textbf{} & \textbf{Present} & \textbf{Not Present} \\
+        \hline
+        \hline
+        Emoji     & 4805 (76.6\%)    & 23952 (68.0\%)       \\
+        Hashtags  & 2122 (70.5\%)    & 26635 (69.4\%)       \\
+
+    \end{tabular}
+
+    \caption{Number of tweets on the test set with and without emoji and
+    hashtags. The number between parentheses is the proportion of tweets
+    correctly classified.}
+
+    \label{table:emoji_and_hashtags}
+
+\end{table}
+
+Table \ref{table:emoji_and_hashtags} shows the overall effect of hashtags and
+emoji on classification performance. Tweets containing emoji seem to be easier
+for the model to classify than those without. Hashtags also have a positive
+effect on classification performance, however it is less significant. This implies that emoji, and hashtags in a smaller degree, provide
+tweets with a context richer in sentiment information, allowing the model to
+better guess the emotion of the \textit{trigger-word}.
+
+Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on classification performance.
+
+\begin{table}[!h]
+    \centering
+    \scriptsize
+        \begin{tabular}{lr|rc|rc|r}
+
+           \multirow{2}{*}{\textbf{Emoji alias}} & \multirow{2}{*}{\textbf{N}} & \multicolumn{2}{c|}{\textbf{emoji}} & \multicolumn{2}{c|}{\textbf{no-emoji}} &
+            \multirow{2}{*}{$\bf{\Delta}$\textbf{\%}} \\
+            \cline{3-6}
+                                &     & \#   & \% & \# & \% & \\
+          \hline
+          \hline
+          \texttt{mask}         & 163 & 154 & 94.48 & 134 & 82.21 & - 12.27 \\
+          \texttt{two\_hearts}  & 87  & 81  & 93.10 & 77  & 88.51 & - 4.59  \\
+          \texttt{heart\_eyes}  & 122 & 109 & 89.34 & 103 & 84.43 & - 4.91  \\
+          \texttt{heart}        & 267 & 237 & 88.76 & 235 & 88.01 & - 0.75  \\
+          \texttt{rage}         & 92  & 78  & 84.78 & 66  & 71.74 & - 13.04 \\
+          \texttt{cry}          & 116 & 97  & 83.62 & 83  & 71.55 & - 12.07 \\
+          \texttt{sob}          & 490 & 363 & 74.08 & 345 & 70.41 & - 3.67  \\
+          \texttt{unamused}     & 167 & 121 & 72.46 & 116 & 69.46 & - 3.00  \\
+          \texttt{weary}        & 204 & 140 & 68.63 & 139 & 68.14 & - 0.49  \\
+          \texttt{joy}          & 978 & 649 & 66.36 & 629 & 64.31 & - 2.05  \\
+          \texttt{sweat\_smile} & 111 & 73  & 65.77 & 75  & 67.57 & 1.80 \\
+          \texttt{confused}     & 77  & 46  & 59.74 & 48  & 62.34 & 2.60 \\
+
+        \end{tabular}
+
+    \caption{Fine grained performance on tweets containing emoji, and the effect of removing them.
+\textbf{N} is the total number of tweets containing the listed emoji, \textbf{\#} and \textbf{\%} the number and percentage of correctly-classified tweets respectively, and $\bf{\Delta}$\textbf{\%} the variation of test accuracy when removing the emoji from the tweets.}
+
+    \label{table:emoji_fine_grained}
+
+\end{table}
+
+It is clear some emoji strongly contribute to improving prediction quality.
+The most interesting ones are \texttt{mask}, \texttt{rage}, and \texttt{cry}, which significantly increase accuracy.
+Further, contrary to intuition, the \texttt{sob} emoji contributes less than \texttt{cry}, despite representing a stronger emotion.
+This is probably due to \texttt{sob} being used for depicting a wider spectrum of emotions.
+
+Finally, not all emoji are beneficial for this task.
+When removing \texttt{sweat\_smile} and \texttt{confused} accuracy increased, probably because they represent emotions other than the ones being predicted.
+%From observing Table \ref{table:emoji_label} it seems likely that tweets containing \texttt{mask} would be highly predictive of the \texttt{disgust} class, since more than $92\%$ of the tweets containing such emoji belong to that class, however it less clear why \texttt{rage}, and \texttt{cry} contribute similarly, despite being more evenly distributed among classes.
+
+%This means that they are highly representative of the the class the tweet belongs to, and would be good choices for discriminating new tweets in case we wanted to expand the dataset.
+
+%For example, by observing Table \ref{table:emoji_label} we can infer that  \texttt{mask}, \texttt{rage}, and \texttt{cry}   \texttt{disgust}, \texttt{rage}, and \texttt{sad} classes respectively.
+
+\subsection{Effect of the Amount of Training Data}
+
+As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data in which our model was trained consistently increased validation accuracy and validation macro F1 score.
+The trend shows that having more training data would likely further increase the model's performance.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=\columnwidth]{images/acc_vs_tdp_variation.eps}
+    \caption{Validation accuracy and macro F1 score vs. proportion of the training data.}
+    \label{fig:data_amt_vs_acc}
+\end{figure}
 
 \section{Conclusions and Future Work}
+
+We described the model that got second place in the WASSA 2018 Implicit Emotion Shared Task.
+Despite its simplicity, and low amount of dependencies on libraries and external features, it performed almost as well as the system that obtained the first place.
+
+Our ablation study revealed that our hyperparameters were indeed quite well-tuned for the task, which agrees with the good results obtained in the official submission.
+However, the ablation study also showed that increased performance can be obtained by incorporating POS embeddings as additional inputs.
+Further experiments are required to accurately measure the impact that this additional input may have on the results.
+We also think the performance can be boosted by making the architecture more complex, concretely, by using a BiLSTM with multiple layers and skip connections in a way akin to \cite{peters2018deep}, or by making the fully-connected network bigger and deeper.
+
+Finally, our studies regarding the importance of hashtags and emoji in the classification showed that both of them seem to contribute significantly to the performance, although in different measures.
 

### 2018-08-27 paper/paper.tex
Update paper to latest version

@@
 
-    Furthermore, we tokenized the text using a variation of the
-    \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}}
-    script, a Python port of the original \texttt{Twokenize.java}
-    \cite{gimpel-EtAl:2011:ACL-HLT2011}. Concretely, we created an emoji-aware
-    version of it by incorporating knowledge from an emoji
-    database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}}
-    which we slightly modified for avoiding conflict with emoji sharing unicode
-    codes with common glyphs used in Twitter,\footnote{For example, the hashtag
-    emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3},
-which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and for making it compatible with Python 3.
+    Furthermore, we tokenized the text using a variation of the \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}} script, a Python port of the original \texttt{Twokenize.java} \cite{gimpel-EtAl:2011:ACL-HLT2011}.
+Concretely, we created an emoji-aware version of it by incorporating knowledge from an emoji database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}} which we slightly modified for avoiding conflict with emoji sharing unicode codes with common glyphs used in Twitter,\footnote{For example, the hashtag emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3}, which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and for making it compatible with Python 3.
 % \begin{itemize}

### 2018-08-27 paper/paper.tex
Add images for paper, small fixes

@@
 \documentclass[11pt,a4paper]{article}
+\pdfoutput=1
 \usepackage[hyperref]{emnlp2018}
@@
   The University of Tokyo\\
-  {jorge, emarrese, matsuo@weblab.t.u-tokyo.ac.jp} \\
+  \{jorge, emarrese, matsuo\}@weblab.t.u-tokyo.ac.jp \\
 }
@@
 We hypothesize this is due to tweets being too short for needing such a rich representation.
-Also, the size of the concatenated vector was $4096\times3=12,288$, which probably could not be properly exploited by the 512-dimensional fully-connected layer.
+Also, the size of the concatenated vector was $4096\times3=12,288$, which probably could not be properly exploited by the $512$-dimensional fully-connected layer.
 
@@
     hashtags. The number between parentheses is the proportion of tweets
-    correctly classified.}
+    classified correctly.}
 

### 2018-08-28 paper/paper.tex
Update some captions and figures in paper

@@
 \usepackage{siunitx}
+\usepackage{caption}
+\usepackage{blindtext}
 \usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
@@
     \includegraphics[width=\columnwidth]{images/best_ensembles.eps}
-\caption{Best Macro F1 vs. Number of Ensembled Models.}
+\caption{Effect of the number of ensembled models on validation performance.}
 \label{fig:best_ensembles}
@@
     \end{tabular}
+
     \caption{Ablation study results.}
+    \vspace{-0.4cm}
+    \caption*{
+        \footnotesize Accuracies were obtained from the
+        validation dataset. Each model was trained with the same random seed and
+        hyperparameters, save for the one listed. ``No emoji'' is the same model
+        trained on the training dataset with no emoji, ``No ELMo'' corresponds
+        to having switched the ELMo word encoding layer with a simple
+        pre-trained GloVe embedding lookup table, and ``Concat Pooling''
+        obtained sentence representations by using the pooling method described
+        by \citet{howard2018universal}. ``LSTM hidden'' corresponds to the the
+        hidden dimension of the BiLSTM, ``POS emb dim'' to the dimension of the
+        part-of-speech embeddings, and ``SGD optim lr'' to the learning rate
+        used while optimizing with the schedule described by \citet{conneau2017supervised}.}
+
     \label{table:ablation}
@@
 \end{table}
-
-Accuracies were obtained from the
-validation dataset.
-Each model was trained with the same random seed and hyperparameters, save for the one listed.
-``No emoji'' is the same model trained on the training dataset with no emoji, ``No ELMo'' corresponds to having switched the ELMo word encoding layer with a simple pre-trained GloVe embedding lookup, and ``Concat Pooling'' obtained sentence representations by using the pooling method described by \citet{howard2018universal}.
-``LSTM hidden'' corresponds to the the hidden dimension of the BiLSTM, ``POS emb dim'' to the dimension of the part-of-speech embeddings, and ``SGD optim lr'' to the learning rate used while optimizing with the schedule described by \citet{conneau2017supervised}.
 
@@
 
-Finally, Figure \ref{fig:dropouts} shows the effect of using different dropout probabilities.
-We can see that having higher dropout after the word-representation layer and the fully-connected network's hidden layer, while having a low dropout after the sentence encoding layer yielded better results overall.
+Finally, Figure \ref{fig:dropouts} shows the effect of using different dropout
+probabilities. We can see that having higher dropout after the
+word-representation layer and the fully-connected network's hidden layer, while
+having a low dropout after the sentence encoding layer yielded better results
+overall.
 
@@
     \includegraphics[width=\columnwidth]{images/dropout_table.pdf}
-\caption{Dropout Variations.
-Rows correspond to the dropout applied both after the ELMo layer (word encoding layer) and after the fully-connected network's hidden layer, while columns correspond to the dropout applied after the max-pooling operation (sentence encoding layer.}
+
+    \caption{Dropout Ablation. }
+    \vspace{-0.4cm}
+
+    \caption*{
+        \footnotesize Rows correspond to the dropout applied both after
+        the ELMo layer (word encoding layer) and after the fully-connected network's
+        hidden layer, while columns correspond to the dropout applied after the
+        max-pooling operation (sentence encoding layer.)}
+
 \label{fig:dropouts}
@@
 
-    \caption{Fine grained performance on tweets containing emoji, and the effect of removing them.
-\textbf{N} is the total number of tweets containing the listed emoji, \textbf{\#} and \textbf{\%} the number and percentage of correctly-classified tweets respectively, and $\bf{\Delta}$\textbf{\%} the variation of test accuracy when removing the emoji from the tweets.}
+    \caption{Fine grained performance on tweets containing emoji, and the effect
+of removing them.} 
+    \vspace{-0.4cm}
+    \caption*{
+        \footnotesize \textbf{N} is the total number of tweets containing the listed emoji,
+        \textbf{\#} and \textbf{\%} the number and percentage of correctly-classified
+        tweets respectively, and $\bf{\Delta}$\textbf{\%} the variation of test accuracy
+        when removing the emoji from the tweets.}
 

### 2018-08-28 src/models/transformer.py
Remove matplotlib dependencies

@@ -12,8 +12,6 @@ import torch.nn as nn
 import torch.nn.functional as F
 
 import numpy as np
-# import matplotlib.pyplot as plt
-# import seaborn
 
 import colored_traceback
 colored_traceback.add_hook(always=True)

### 2018-08-28 src/optim/optim.py

@@ -113,14 +113,3 @@ class OptimWithDecay(object):
 
         self._makeOptimizer()
         return updated, self.lr
-
-
-if __name__ == "__main__":
-
-    import numpy as np
-    import matplotlib.pyplot as plt
-
-    max_step = 2000
-    plt.plot(np.arange(1, max_step), [slanted_triangular_lr(step, max_step) for step in range(1, max_step)])
-    plt.show()
-

### 2018-08-28 src/utils/logger.py

@@ -9,10 +9,6 @@ from datetime import datetime
 # shouldn't do this!! the point is for this not to be library specific
 import torch
 
-import matplotlib
-matplotlib.use('Agg')
-import matplotlib.pyplot as plt
-
 from .io import (write_output)
 from .. import config
 
@@ -204,13 +200,6 @@ class Logger(object):
         torch.save(obj, savepath)
         if progress_bar:
             progress_bar.write(f'File saved in {savepath}')
-        return
-
-    def save_image(self, numpy_matrix, filename='image', path_override=None):
-        plt.imshow(numpy_matrix)
-        savepath = os.path.join(self.run_savepath, filename + '.png')
-        savepath = path_override if path_override else savepath
-        plt.savefig(savepath)
         return
 
     def write_output_details(self, corpus, pred_label_ids,

### 2018-08-28 src/config.py
Update README with easier way to install codebase

@@ -1,7 +1,7 @@
 import os
 
 HOME_DIR = os.environ['HOME']
-ALLENNLP_PATH = os.path.join(HOME_DIR, 'repos/allennlp')
+ALLENNLP_PATH = 'allennlp/'
 
 DATA_PATH = 'data/'
 CACHE_PATH = '.cache'

### 2018-08-28 install.sh
Add install script, update README.md

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+conda install -y pytorch=0.4.0 cuda80 -c pytorch
+conda install -y scikit-learn
+pip install -r requirements.txt
+
+git clone https://github.com/allenai/allennlp.git
+cd allennlp
+# Using exactly the same version we used for development
+git checkout ac2e0b9b6e4668984ebd8c05578d9f4894e94bee
+INSTALL_TEST_REQUIREMENTS=false scripts/install_requirements.sh

### 2018-08-29 get_data.sh
Tidy up preprocessing script, separate logic for obtaining POS tags
Update README, delete unused file

@@ -14,6 +14,5 @@ gunzip train-v3.csv.gz
 gunzip trial-v3.csv.gz
 gunzip trial-v3.labels.gz
 gunzip test.csv.gz
-gunzip test.labels.gz
 
 cd ..

### 2018-08-29 get_pos.sh

@@ -1,0 +1,58 @@
+#!/bin/bash
+
+PREPROCESSED_DATA_DIR="data/preprocessed"
+
+CLEAN_TRAIN_DATA_PATH="$PREPROCESSED_DATA_DIR/train.csv"
+TRAIN_LABELS_PATH="$PREPROCESSED_DATA_DIR/train_labels.csv"
+
+CLEAN_DEV_DATA_PATH="$PREPROCESSED_DATA_DIR/dev.csv"
+DEV_LABELS_PATH="$PREPROCESSED_DATA_DIR/dev_labels.csv"
+
+CLEAN_TEST_DATA_PATH="$PREPROCESSED_DATA_DIR/test.csv"
+TEST_LABELS_PATH="$PREPROCESSED_DATA_DIR/test_labels.csv"
+
+
+#######################################################################
+#                                TRAIN                                #
+#######################################################################
+
+
+echo "Getting POS tags for $CLEAN_TRAIN_DATA_PATH with TwiboParser"
+
+# this script will generate the file $CLEAN_TRAIN_DATA_PATH.tagged.pre
+./utils/run_postagger.sh $CLEAN_TRAIN_DATA_PATH
+
+cat $CLEAN_TRAIN_DATA_PATH.tagged.pre | ./utils/postprocess_postag_file.py > $CLEAN_TRAIN_DATA_PATH.tagged
+rm $CLEAN_TRAIN_DATA_PATH.tagged.pre
+
+# this script will generate two files:
+# - $CLEAN_TRAIN_DATA_PATH.tagged.tokens
+# - $CLEAN_TRAIN_DATA_PATH.tagged.pos
+./utils/parse_twibo_output.py $CLEAN_TRAIN_DATA_PATH.tagged
+
+echo
+#######################################################################
+#                                 DEV                                 #
+#######################################################################
+echo "Getting POS tags for $CLEAN_DEV_DATA_PATH with TwiboParser"
+
+./utils/run_postagger.sh $CLEAN_DEV_DATA_PATH
+
+cat $CLEAN_DEV_DATA_PATH.tagged.pre | ./utils/postprocess_postag_file.py > $CLEAN_DEV_DATA_PATH.tagged
+rm $CLEAN_DEV_DATA_PATH.tagged.pre
+
+python ./utils/parse_twibo_output.py $CLEAN_DEV_DATA_PATH.tagged
+
+echo
+#######################################################################
[…]

### 2018-08-29 preprocess.sh

@@ -31,111 +31,108 @@ CLEAN_TRAIN_NO_EMOJIS_DATA_PATH="$PREPROCESSED_DATA_DIR/train_no_emojis.csv"
 CLEAN_DEV_NO_EMOJIS_DATA_PATH="$PREPROCESSED_DATA_DIR/dev_no_emojis.csv"
 CLEAN_TEST_NO_EMOJIS_DATA_PATH="$PREPROCESSED_DATA_DIR/test_no_emojis.csv"
 
+################################################################################
+#                           No-Triggerword data 
+################################################################################
+CLEAN_TRAIN_NO_TRIGGERWORD_DATA_PATH="$PREPROCESSED_DATA_DIR/train_no_triggerword.csv"
+CLEAN_DEV_NO_TRIGGERWORD_DATA_PATH="$PREPROCESSED_DATA_DIR/dev_no_triggerword.csv"
+CLEAN_TEST_NO_TRIGGERWORD_DATA_PATH="$PREPROCESSED_DATA_DIR/test_no_triggerword.csv"
 
-mkdir $PREPROCESSED_DATA_DIR
-echo "Created dir $PREPROCESSED_DATA_DIR"
+function preprocess () {
+    INFILE=$1
+    OUTFILE=$2
+
+    # The first awk call filters out those examples that do not contain
+    # [#TRIGGERWORD#] The second one replaces keywords with a more standard
+    # notation and adds spaces around them to make the job easier for the
+    # tokenizer
+
+    cat $INFILE | awk '{if ( $0 ~ /#TRIGGERWORD#/ ) {print $0}}' \
+                | awk '{$1="";
+                        gsub("\\[#TRIGGERWORD#\\]", " __TRIGGERWORD__ ", $0); 
+                        gsub("@USERNAME", " __USERNAME__ ", $0); 
+                        gsub("\\[NEWLINE\\]", " __NEWLINE__ ", $0); 
+                        gsub("http://url.removed", " __URL__ ", $0);
+                        print $0}' \
+                | ./utils/twokenize.py > $OUTFILE
+}
+
+if [ ! -d $PREPROCESSED_DATA_DIR ]; then
+    mkdir $PREPROCESSED_DATA_DIR
+    echo "Created dir $PREPROCESSED_DATA_DIR"
+fi
 
 echo
+
[…]

### 2018-08-29 src/seq_utils/data_manager.py

@@ -1,220 +1,0 @@
-from pad import pad1d, pad2d
-
-
-def map_sequence(seq, sequence_map, unk_item_id):
-    """ Transform a splitted sequence of items into another sequence of items
-        according to the rules encoded in the dict item2id
-       seq: iterable
-       sequence_map: dict
-       unk_item_id: int"""
-
-    item_ids = []
-    for item in seq:
-        item_id = sequence_map.get(item, unk_item_id)
-        item_ids.append(item_id)
-    return item_ids
-
-
-def map_sequences(sequences, sequence_map, unk_item_id):
-    """Transform a list of sequences into another one, according to
-       the rules encoded in sequence map"""
-
-    mapped_sequences = []
-    for seq in sequences:
-        mapped_sequence = map_sequence(seq, sequence_map, unk_item_id)
-        mapped_sequences.append(mapped_sequence)
-    return mapped_sequences
-
-
-def split_map_sequence(seq, sequence_map, unk_item_id, seq_splitter):
-    """ Transform a sequence of items into another sequence of items
-        according to the rules encoded in the dict item2id.
-
-        Example usage: mapping words into their corresponding ids
-
-       seq: iterable
-       sequence_map: dict
-       unk_item_id: int
-       seq_splitter: function"""
-
-    splitted_seq = seq_splitter(seq)
-    item_ids = map_sequence(splitted_seq, sequence_map, unk_item_id)
-    return item_ids
-
-
-def split_map_sequences(sequences, sequence_map, unk_item_id, seq_splitter):
-    """Split the sequences and then transform them into the items specified
-       by sequence_map"""
-
-    splitted_seqs = [seq_splitter(seq) for seq in sequences]
-    splitted_mapped_seqs = map_sequences(splitted_seqs, sequence_map,
-                                         unk_item_id)
-    return splitted_mapped_seqs
-
-
-def split_map_pad_sequences(sequences, sequence_map, unk_item_id, pad_id,
-                            seq_splitter):
-    """Split, transform (map) and pad a batch of sequences
-
-       return the padded and mapped sequences, along with the original lengths
-       and a mask indicating the real item positions, as opposed to the
[…]

### 2018-08-29 get_data.sh
Get pre-trained embeddings and ELMo weights in get_data.sh script

@@ -3,6 +3,10 @@
 USER=$1
 
 cd data
+################################################################################
+#                           TRAIN / DEV / TEST data                            #
+################################################################################
+
 wget --user=$USER --ask-password http://implicitemotions.wassa2018.com/data/protected/train-v3.csv.gz
 wget http://implicitemotions.wassa2018.com/data/unprotected/trial-v3.csv.gz
 wget http://implicitemotions.wassa2018.com/data/unprotected/trial-v3.labels.gz
@@ -15,4 +19,18 @@ gunzip trial-v3.csv.gz
 gunzip trial-v3.labels.gz
 gunzip test.csv.gz
 
-cd ..
+mkdir word_embeddings
+cd word_embeddings
+################################################################################
+#                         Pre-trained GloVe embeddings                         #
+################################################################################
+wget http://nlp.stanford.edu/data/glove.840B.300d.zip
+unzip glove.840B.300d.zip
+
+################################################################################
+#                           Pre-trained ELMo Weights                           #
+################################################################################
+wget https://s3-us-west-2.amazonaws.com/allennlp/models/elmo/2x4096_512_2048cnn_2xhighway/elmo_2x4096_512_2048cnn_2xhighway_options.json
+wget https://s3-us-west-2.amazonaws.com/allennlp/models/elmo/2x4096_512_2048cnn_2xhighway/elmo_2x4096_512_2048cnn_2xhighway_weights.hdf5
+
+cd ../..

### 2018-08-29 install.sh
Small fixes in scripts, better prediction filenames
Save hyperparams.json in pretty format

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 conda install -y pytorch=0.4.0 cuda80 -c pytorch
-conda install -y scikit-learn
+conda install -y scikit-learn unzip
 pip install -r requirements.txt
 
 git clone https://github.com/allenai/allennlp.git

### 2018-08-29 preprocess.sh

@@ -130,9 +130,7 @@ cat $TEST_DATA_PATH | awk '{if ( $0 ~ /#TRIGGERWORD#/ ) {print $1}}' > $TEST_LAB
 echo "Created $CLEAN_TEST_DATA_PATH and $TEST_LABELS_PATH"
 echo
 
-echo
-
-echo "Removing emojis from test"
+echo "Creating Test dataset with no emoji"
 cat $CLEAN_TEST_DATA_PATH | ./utils/remove_emojis.py > $CLEAN_TEST_NO_EMOJIS_DATA_PATH
 echo "Created $CLEAN_TEST_NO_EMOJIS_DATA_PATH"
 

### 2018-08-29 run.py

@@ -134,6 +134,9 @@ def main():
     if hp.write_mode != 'NONE':
         logger.write_hyperparams()
 
+    print(f'Running experiment {logger.model_mash}. Hyperparameters and '
+          f'checkpoints will be saved in {logger.run_savepath}')
+
     torch.manual_seed(hp.seed)
     torch.cuda.manual_seed_all(hp.seed)  # silently ignored if there are no GPUs
 
@@ -309,21 +312,17 @@ def main():
             corpus.train_batches.shuffle_examples()
             eval_dict = trainer.evaluate(corpus.dev_batches, epoch, writer)
 
-            if hp.update_learning_rate:
-                if hp.model != 'transformer':
-                    optim_updated, new_lr = trainer.optimizer.updt_lr_accuracy(epoch, eval_dict['accuracy'])
-                    lr_threshold = 1e-5
-                    if new_lr < lr_threshold:
-                        tqdm.write(f'Learning rate smaller than {lr_threshold}, '
-                                   f'stopping.')
-                        break
-                    if optim_updated:
-                        tqdm.write(f'Learning rate decayed to {new_lr}')
+            if hp.update_learning_rate and hp.model != 'transformer':
+                optim_updated, new_lr = trainer.optimizer.updt_lr_accuracy(epoch, eval_dict['accuracy'])
 
-            # elif hp.update_learning_rate_nie:
-            #     optim_updated, new_lr = trainer.optimizer.update_learning_rate_nie(epoch)
-            #     if optim_updated:
-            #         tqdm.write(f'Learning rate decayed to {new_lr}')
+                #  TODO: lr_threshold shouldn't be hardcoded
+                lr_threshold = 1e-5
+                if new_lr < lr_threshold:
+                    tqdm.write(f'Learning rate smaller than {lr_threshold}, '
+                               f'stopping.')
+                    break
+                if optim_updated:
+                    tqdm.write(f'Learning rate decayed to {new_lr}')
 
             accuracy = eval_dict['accuracy']
             if not best_accuracy or accuracy > best_accuracy:
@@ -334,12 +333,12 @@ def main():
                 if hp.write_mode != 'NONE':
                     probs = np_softmax(eval_dict['output'])
                     probs_filepath = os.path.join(logger.run_savepath,
-                                                  'best_eval_probs.csv')
+                                                  'best_dev_probabilities.csv')
                     np.savetxt(probs_filepath, probs,
                                delimiter=',', fmt='%.8f')
 
[…]

### 2018-08-29 src/utils/logger.py

@@ -105,7 +105,6 @@ class Logger(object):
         self.args['commit'] = get_commit_hash()
 
         if not experiment_hash:
-            # For hash based instead of datetime based monitoring
             hashargs = {arg: val for arg, val in self.args.items()
                         if arg not in ['datetime', 'auto_submission',
                                        'no_tensorboard', 'write_mode',
@@ -128,7 +127,7 @@ class Logger(object):
             hyperparam_file = os.path.join(self.run_savepath,
                                            'hyperparams.json')
             with open(hyperparam_file, 'w') as f:
-                f.write(json.dumps(self.args))
+                f.write(json.dumps(self.args, indent=2, separators=(',', ': ')))
 
         if write_mode == 'DATABASE' or write_mode == 'BOTH':
             db = dt.connect(config.DATABASE_CONNECTION_STRING)

### 2018-08-29 run.py
Fix typo in run.py

@@ -134,7 +134,7 @@ def main():
     if hp.write_mode != 'NONE':
         logger.write_hyperparams()
 
-    print(f'Running experiment {logger.model_mash}. Hyperparameters and '
+    print(f'Running experiment {logger.model_hash}. Hyperparameters and '
           f'checkpoints will be saved in {logger.run_savepath}')
 
     torch.manual_seed(hp.seed)

### 2018-08-29 run.py
Remove old unused hyperparam (nie learning rate decay)

@@ -75,8 +75,6 @@ arg_parser.add_argument('--dropout', type=float, default=0.1,
 arg_parser.add_argument('--model_hash', type=str, default=None,
                         help='Hash of the model to load, can be a partial hash')
 
-arg_parser.add_argument('--update_learning_rate_nie', '-ulrn', action='store_true')
-
 arg_parser.add_argument('--word_encoding_method', '-wem', type=str, default="embed",
                         choices=WordEncodingLayer.WORD_ENCODING_METHODS,
                         help='How to obtain word representations')
@@ -118,7 +116,6 @@ arg_parser.add_argument("--spreadsheet", "-ss", action='store_true',
 
 def validate_args(hp):
     """hp: argparser parsed arguments. type: Namespace"""
-    assert not (hp.update_learning_rate and hp.update_learning_rate_nie)
 
     if hp.word_encoding_method == 'char_lstm' and not hp.word_char_aggregation_method:
         raise ValueError(f'Need to pass a word_char_aggregation_method when '

### 2018-08-30 run.py
Make best hyperparams default, tidy up run.py file

@@ -33,15 +33,14 @@ from base_args import base_parser, CustomArgumentParser
 colored_traceback.add_hook(always=True)
 
 
-base_parser.description = 'PyTorch MultiNLI Inner Attention Classifier'
 arg_parser = CustomArgumentParser(parents=[base_parser],
-                                  description='PyTorch MultiNLI')
+                                  description='Implicit Emotion Classifier')
 
 arg_parser.add_argument('--model', type=str, default="bilstm",
                         choices=SentenceEncodingLayer.SENTENCE_ENCODING_METHODS,
                         help='Model to use')
 
-arg_parser.add_argument('--corpus', type=str, default="iest",
+arg_parser.add_argument('--corpus', type=str, default="iest_emoji",
                         choices=list(config.corpora_dict.keys()),
                         help='Name of the corpus to use.')
 
@@ -58,24 +57,27 @@ arg_parser.add_argument('--sent_enc_layers', type=int, default=1,
 arg_parser.add_argument('--force_reload', action='store_true',
                         help='Whether to reload pickles or not (makes the '
                         'process slower, but ensures data coherence)')
-arg_parser.add_argument('--char_emb_dim', '-ced', type=int, default=50,
+
+arg_parser.add_argument('--char_emb_dim', '-ced', type=int, default=200,
                         help='Char embedding dimension')
+
 arg_parser.add_argument('--pooling_method', type=str, default='max',
                         choices=PoolingLayer.POOLING_METHODS,
                         help='Pooling scheme to use as raw sentence '
                              'representation method.')
 
-arg_parser.add_argument('--sent_enc_dropout', type=float, default=0.0,
+arg_parser.add_argument('--sent_enc_dropout', type=float, default=0.2,
                         help='Dropout between sentence encoding lstm layers. '
                              'and after the sent enc lstm. 0 means no dropout.')
 
-arg_parser.add_argument('--dropout', type=float, default=0.1,
-                        help='Dropout applied to layers and final MLP. 0 means no dropout.')
+arg_parser.add_argument('--dropout', type=float, default=0.5,
+                        help='Dropout applied to words representations and '
+                        'final MLP. 0 means no dropout.')
 
 arg_parser.add_argument('--model_hash', type=str, default=None,
                         help='Hash of the model to load, can be a partial hash')
 
-arg_parser.add_argument('--word_encoding_method', '-wem', type=str, default="embed",
+arg_parser.add_argument('--word_encoding_method', '-wem', type=str, default="elmo",
[…]

### 2018-08-30 src/utils/logger.py
Avoid printing fatal error when current dir is not a git repo

@@ -48,7 +48,8 @@ def get_server_name():
 
 def get_commit_hash():
     try:
-        output = subprocess.check_output(['git', 'rev-parse', 'HEAD']).strip()
+        output = subprocess.check_output(['git', 'rev-parse', 'HEAD'],
+                                         stderr=subprocess.PIPE).strip()
         return output.decode()
     except subprocess.CalledProcessError:
         print('Current project is not a git repo, omitting commit hash.')

### 2018-08-30 run.py
Fix default sentence encoder dropout

@@ -66,7 +66,7 @@ arg_parser.add_argument('--pooling_method', type=str, default='max',
                         help='Pooling scheme to use as raw sentence '
                              'representation method.')
 
-arg_parser.add_argument('--sent_enc_dropout', type=float, default=0.2,
+arg_parser.add_argument('--sent_enc_dropout', type=float, default=0.1,
                         help='Dropout between sentence encoding lstm layers. '
                              'and after the sent enc lstm. 0 means no dropout.')
 

### 2018-08-30 src/models/iest.py
Fix lstm layer dropouts at 0, store sentence reprs, make pos optional
POS tags are now optional in the call to the encode method of the
IESTClassifier.

Sentence representations are stored for later analyses

@@ -283,11 +283,18 @@ class BLSTMEncoder(nn.Module):
                  use_cuda=True):
         super(BLSTMEncoder, self).__init__()
         self.embedding_dim = embedding_dim
-        self.hidden_size = hidden_sizes  # sizes in plural for compatibility
+
+        # hidden_sizes in plural for compatibility with StackedShortcutLSTM
+        self.hidden_size = hidden_sizes
+
         self.num_layers = num_layers
         self.bidirectional = bidirectional
         self.num_dirs = 2 if bidirectional else 1
-        self.dropout = dropout
+
+        # FIXME: Layer dropouts fixed at 0 for now
+        # self.dropout = dropout
+        self.dropout = 0.0
+
         self.batch_first = batch_first
         self.out_dim = self.hidden_size * self.num_dirs
 
@@ -296,7 +303,7 @@ class BLSTMEncoder(nn.Module):
                                 dropout=self.dropout)
 
     def is_cuda(self):
-        # FIXME: Avoid calls to data()
+        # FIXME: Avoid calls to data() as of pytorch 0.4.0
         # either all weights are on cpu or they are on gpu
         return 'cuda' in str(type(self.enc_lstm.bias_hh_l0.data))
 
@@ -455,9 +462,9 @@ class IESTClassifier(nn.Module):
             nn.Linear(512, self.num_classes)
         )
 
-    def encode(self, batch, char_batch, pos_batch,
+    def encode(self, batch, char_batch,
                sent_lengths, word_lengths,
-               masks=None, char_masks=None, embed_words=True,
+               pos_batch=None, masks=None, char_masks=None, embed_words=True,
                raw_sequences=None):
         """ Encode a batch of ids into a sentence representation.
 
@@ -552,18 +559,18 @@ class IESTClassifier(nn.Module):
                                 self.use_cuda,
                                 requires_grad=False)
 
-        sent_vec = self.encode(sequences,
-                               char_batch=char_sequences,
-                               pos_batch=pos_sequences,
-                               sent_lengths=sent_lengths,
-                               word_lengths=word_lengths,
-                               masks=masks,
-                               char_masks=char_masks,
-                               raw_sequences=raw_sequences)
+        self.sent_vec = self.encode(sequences,
+                                    char_batch=char_sequences,
+                                    pos_batch=pos_sequences,
+                                    sent_lengths=sent_lengths,
+                                    word_lengths=word_lengths,
+                                    masks=masks,
[…]

### 2018-08-30 run.py
Specify lstm layer dropout in CLI
Default is 0.0

@@ -66,13 +66,17 @@ arg_parser.add_argument('--pooling_method', type=str, default='max',
                         help='Pooling scheme to use as raw sentence '
                              'representation method.')
 
-arg_parser.add_argument('--sent_enc_dropout', type=float, default=0.1,
-                        help='Dropout between sentence encoding lstm layers. '
-                             'and after the sent enc lstm. 0 means no dropout.')
-
 arg_parser.add_argument('--dropout', type=float, default=0.5,
                         help='Dropout applied to words representations and '
                         'final MLP. 0 means no dropout.')
+
+arg_parser.add_argument('--lstm_layer_dropout', type=float, default=0.0,
+                        help='Dropout between sentence encoding lstm layers. '
+                             '0 means no dropout.')
+
+arg_parser.add_argument('--sent_enc_dropout', type=float, default=0.1,
+                        help='Dropout after the sentence encoding lstm. '
+                        '0 means no dropout.')
 
 arg_parser.add_argument('--model_hash', type=str, default=None,
                         help='Hash of the model to load, can be a partial hash')
@@ -215,6 +219,7 @@ def main():
             pooling_method=hp.pooling_method,
             batch_first=True,
             dropout=hp.dropout,
+            lstm_layer_dropout=hp.lstm_layer_dropout,
             sent_enc_dropout=hp.sent_enc_dropout,
             sent_enc_layers=hp.sent_enc_layers
         )

### 2018-08-30 src/models/iest.py

@@ -395,6 +395,7 @@ class IESTClassifier(nn.Module):
                  pooling_method='max',
                  batch_first=True,
                  dropout=0.0,
+                 lstm_layer_dropout=0.0,
                  sent_enc_dropout=0.0,
                  use_cuda=True):
 
@@ -403,6 +404,7 @@ class IESTClassifier(nn.Module):
         self.batch_size = batch_size
         self.batch_first = batch_first
         self.dropout = dropout
+        self.lstm_layer_dropout = lstm_layer_dropout
         self.sent_enc_dropout = sent_enc_dropout
 
         self.use_cuda = use_cuda
@@ -443,7 +445,7 @@ class IESTClassifier(nn.Module):
             num_layers=self.sent_enc_layers,
             batch_first=self.batch_first,
             use_cuda=self.use_cuda,
-            dropout=self.sent_enc_dropout
+            dropout=self.lstm_layer_dropout
         )
 
         self.sent_dropout = nn.Dropout(self.sent_enc_dropout)

### 2018-08-30 base_args.py
Set default hyperparams in base_args, add some comments

@@ -95,9 +95,9 @@ misc_args.add_argument('--no_cuda', action='store_true',
                        help='Force the use of the cpu even if a gpu is '
                             'available')
 
-misc_args.add_argument('--log_interval', type=int, default=200, metavar='',
+misc_args.add_argument('--log_interval', type=int, default=50, metavar='',
                        help='Number of iterations between training loss '
                             'loggings')
 
-misc_args.add_argument('--seed', type=int, default=42, metavar='',
+misc_args.add_argument('--seed', type=int, default=43, metavar='',
                        help='Random seed to be used by torch initializations')

### 2018-08-30 src/layers/elmo.py

@@ -18,6 +18,9 @@ class ElmoWordEncodingLayer(object):
         self._embedder = Elmo(config.ELMO_OPTIONS, config.ELMO_WEIGHTS,
                               num_output_representations=1, **kwargs)
         self._embedder = self._embedder.cuda()
+        # We know the output of ELMo with pre-trained weigths is of size 1024.
+        # This would likely be different if we initialized `Elmo` with a custom
+        # `module`, but we didn't test this, so for now it will be hardcoded.
         self.embedding_dim = 1024
 
     def __call__(self, *args):
@@ -25,7 +28,9 @@ class ElmoWordEncodingLayer(object):
 
         # -1 is the raw_sequences element passed in the encode function of the
         # IESTClassifier
+        # TODO: make this less hacky
         sents = args[-1]
+
         char_ids = batch_to_ids(sents)
         char_ids = to_var(char_ids,
                           use_cuda=True,

### 2018-08-30 src/utils/logger.py

@@ -48,6 +48,9 @@ def get_server_name():
 
 def get_commit_hash():
     try:
+        # We pipe stderr because when the directory is not part of a git repo
+        # it prints a fatal error, which we handle here. We don't want to print
+        # that.
         output = subprocess.check_output(['git', 'rev-parse', 'HEAD'],
                                          stderr=subprocess.PIPE).strip()
         return output.decode()

### 2018-09-01 paper/paper.tex
[paper] Add acknowledgements, wrap lines, small fixes

@@
 \begin{abstract}
-In this paper we describe our system designed for the WASSA 2018 Implicit Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 26 teams with a test macro F1 score of $0.710$.
-The system is composed of a single pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory Network BiLSTM for enriching word representations with context, a max-pooling operation for creating sentence representations from said word vectors, and a Dense Layer for projecting the sentence representations into label space.
-Our official submission was obtained by ensembling 6 of these models initialized with different random seeds.
+
+In this paper we describe our system designed for the WASSA 2018 Implicit
+Emotion Shared Task (IEST), which obtained 2$^{\text{nd}}$ place out of 30 teams
+with a test macro F1 score of $0.710$. The system is composed of a single
+pre-trained ELMo layer for encoding words, a Bidirectional Long-Short Memory
+Network BiLSTM for enriching word representations with context, a max-pooling
+operation for creating sentence representations from them, and a
+Dense Layer for projecting the sentence representations into label space. Our
+official submission was obtained by ensembling 6 of these models initialized
+with different random seeds.
+
 \end{abstract}
@@
 
-Although the definition of emotion is still debated among the scientific community, the automatic identification and understanding of human emotions by machines has long been of interest in computer science.
-It has usually been assumed that emotions are triggered by the interpretation of a stimulus event according to its meaning.
+Although the definition of emotion is still debated among the scientific
+community, the automatic identification and understanding of human emotions by
+machines has long been of interest in computer science. It has usually been
+assumed that emotions are triggered by the interpretation of a stimulus event
+according to its meaning.
 
-As language usually reflects the emotional state of an individual, it is natural to study human emotions by understanding how they are reflected in text.
-We see that many words indeed have affect as a core part of their meaning, for example, \textit{dejected} and \textit{wistful} denote some amount of sadness, and are thus associated with sadness.
-On the other hand, some words are associated with affect even though they do not denote affect.
-For example, \textit{failure} and \textit{death} describe concepts that are usually accompanied by sadness and thus they denote some amount of sadness.
-In this context, the task of automatically recognizing emotions from text has recently attracted the attention of researchers in Natural Language Processing.
-This task is usually formalized as the classification of words, phrases, or documents into predefined discrete emotion categories or dimensions.
-Some approaches have aimed at also predicting the degree to which an emotion is expressed in text \cite{wassa_emoint_2017}.
+As language usually reflects the emotional state of an individual, it is natural
+to study human emotions by understanding how they are reflected in text. We see
+that many words indeed have affect as a core part of their meaning, for example,
+\textit{dejected} and \textit{wistful} denote some amount of sadness, and are
+thus associated with sadness. On the other hand, some words are associated with
+affect even though they do not denote affect. For example, \textit{failure} and
+\textit{death} describe concepts that are usually accompanied by sadness and
+thus they denote some amount of sadness. In this context, the task of
+automatically recognizing emotions from text has recently attracted the
+attention of researchers in Natural Language Processing. This task is usually
+formalized as the classification of words, phrases, or documents into predefined
+discrete emotion categories or dimensions. Some approaches have aimed at also
+predicting the degree to which an emotion is expressed in text
+\cite{wassa_emoint_2017}.
 
-In light of this, the \Acronym{WASSA} 2018 Implicit Emotion Shared Task (\Acronym{IEST}) \cite{klinger2018iest} was proposed to help find ways to automatically learn the link between situations and the emotion they trigger.
+In light of this, the \Acronym{WASSA} 2018 Implicit Emotion Shared Task
+(\Acronym{IEST}) \cite{klinger2018iest} was proposed to help find ways to
+automatically learn the link between situations and the emotion they trigger.
 The task consisted in predicting the emotion of a word excluded from a tweet.
-Removed words, or \textit{trigger-words}, included the terms ``sad'', ``happy'', ``disgusted'', ``surprised'', ``angry'', ``afraid'' and their synonyms, and the task was to predict the emotion they conveyed, specifically sadness, joy, disgust, surprise, anger and fear.
+Removed words, or \textit{trigger-words}, included the terms ``sad'', ``happy'',
+``disgusted'', ``surprised'', ``angry'', ``afraid'' and their synonyms, and the
+task was to predict the emotion they conveyed, specifically sadness, joy,
+disgust, surprise, anger and fear.
 
-From a machine learning perspective, this problem can be seen as sentence classification, in which the goal is to classify a sentence, or in particular a tweet, into one of several categories.
-In the case of \Acronym{IEST}, the problem is specially challenging since tweets contain informal language, the heavy usage of emoji, hashtags and username mentions.
+From a machine learning perspective, this problem can be seen as sentence
+classification, in which the goal is to classify a sentence, or in particular a
+tweet, into one of several categories. In the case of \Acronym{IEST}, the
+problem is specially challenging since tweets contain informal language, the
+heavy usage of emoji, hashtags and username mentions.
 
-In this paper we describe our system designed for \Acronym{IEST}, which obtained the second place out of 26 teams.
-Our system did not require manual feature engineering and only minimal use of external data.
-Concretely, our approach is composed of a single pre-trained \Acronym{ELMo} layer for encoding words \cite{peters2018deep}, a Bidirectional Long-Short Memory Network (\Acronym{BiLSTM}) \cite{graves2005framewise, graves2013speech}, for enriching word representations with context, a max-pooling operation for creating sentence representations from said word vectors, and finally a Dense Layer for projecting the sentence representations into label space.
-To the best of our knowledge, our system, which we plan to release, is the first to utilize ELMo for emotion recognition.
+In this paper we describe our system designed for \Acronym{IEST}, which obtained
+the second place out of 26 teams. Our system did not require manual feature
+engineering and only minimal use of external data. Concretely, our approach is
+composed of a single pre-trained \Acronym{ELMo} layer for encoding words
+\cite{peters2018deep}, a Bidirectional Long-Short Memory Network
+(\Acronym{BiLSTM}) \cite{graves2005framewise, graves2013speech}, for enriching
+word representations with context, a max-pooling operation for creating sentence
+representations from said word vectors, and finally a Dense Layer for projecting
+the sentence representations into label space. To the best of our knowledge, our
+system, which we plan to release, is the first to utilize ELMo for emotion
+recognition.
 
@@
 
-As our model is purely character-based, we performed little data preprocessing. Table \ref{table:substitutions} shows the special tokens found in the datasets, and how we substituted them.
+As our model is purely character-based, we performed little data preprocessing.
+Table \ref{table:substitutions} shows the special tokens found in the datasets,
+and how we substituted them.
 
@@
 
-    Furthermore, we tokenized the text using a variation of the \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}} script, a Python port of the original \texttt{Twokenize.java} \cite{gimpel-EtAl:2011:ACL-HLT2011}.
-Concretely, we created an emoji-aware version of it by incorporating knowledge from an emoji database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}} which we slightly modified for avoiding conflict with emoji sharing unicode codes with common glyphs used in Twitter,\footnote{For example, the hashtag emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3}, which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and for making it compatible with Python 3.
+    Furthermore, we tokenized the text using a variation of the
+    \texttt{twokenize.py}\footnote{\tiny\url{https://github.com/myleott/ark-twokenize-py}}
+    script, a Python port of the original \texttt{Twokenize.java}
+    \cite{gimpel-EtAl:2011:ACL-HLT2011}. Concretely, we created an emoji-aware
+    version of it by incorporating knowledge from an emoji
+    database,\footnote{\tiny\url{https://github.com/carpedm20/emoji/blob/e7bff32/emoji/unicode_codes.py}}
+    which we slightly modified for avoiding conflict with emoji sharing unicode
+    codes with common glyphs used in Twitter,\footnote{For example, the hashtag
+    emoji is composed by the unicode code points \texttt{U+23 U+FE0F U+20E3},
+which include \texttt{U+23}, the same code point for the \texttt{\#} glyph.} and
+for making it compatible with Python 3.
+
 % \begin{itemize}
@@
 
-Figure \ref{fig:architecture} summarizes our proposed architecture.
-Our input is based on Embeddings from Language Models (ELMo) by \citet{peters2018deep}.
-These are character-based word representations allowing the model to avoid the ``unknown token'' problem.
-ELMo uses a set of convolutional neural networks to extract features from character embeddings, and builds word vectors from them.
-These are then fed to a multi-layer Bidirectional Language Model (BiLM) which returns context-sensitive vectors for each input word.
+Figure \ref{fig:architecture} summarizes our proposed architecture. Our input is
+based on Embeddings from Language Models (ELMo) by \citet{peters2018deep}. These
+are character-based word representations allowing the model to avoid the
+``unknown token'' problem. ELMo uses a set of convolutional neural networks to
+extract features from character embeddings, and builds word vectors from them.
+These are then fed to a multi-layer Bidirectional Language Model (BiLM) which
+returns context-sensitive vectors for each input word.
 
-We used a single-layer BiLSTM as context fine-tuner \cite{graves2005framewise, graves2013speech}, on top of the ELMo embeddings, and then aggregated the hidden states it returned by using max-pooling, which has been shown to perform well on sentence classification tasks \cite{conneau2017supervised}.
+We used a single-layer BiLSTM as context fine-tuner \cite{graves2005framewise,
+graves2013speech}, on top of the ELMo embeddings, and then aggregated the hidden
+states it returned by using max-pooling, which has been shown to perform well on
+sentence classification tasks \cite{conneau2017supervised}.
 
-Finally, we used a single-layer fully-connected network for projecting the pooled BiLSTM output into a vector corresponding to the label logits for each predicted class.
+Finally, we used a single-layer fully-connected network for projecting the
+pooled BiLSTM output into a vector corresponding to the label logits for each
+predicted class.
 
@@
 
-\hspace{\parindent} \textbf{ELMo Layer}: We used the official AllenNLP implementation of the ELMo model\footnote{\tiny\url{https://allenai.github.io/allennlp-docs/api/allennlp.modules.elmo.html}}, with the official weights pre-trained on the 1 Billion Word Language Model Benchmark, which contains about 800M tokens of news crawl data from WMT 2011 \cite{Chelba2014}.
+\hspace{\parindent} \textbf{ELMo Layer}: We used the official AllenNLP
+implementation of the ELMo
+model\footnote{\tiny\url{https://allenai.github.io/allennlp-docs/api/allennlp.modules.elmo.html}},
+with the official weights pre-trained on the 1 Billion Word Language Model
+Benchmark, which contains about 800M tokens of news crawl data from WMT 2011
+\cite{Chelba2014}.
 
-\textbf{Dimensionalities}: By default the ELMo layer outputs a 2048-dim vector, which we then feed to a BiLSTM resulting in a 4096-dim vector for each word of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read the input in opposite directions and whose outputs are concatenated at the sequence dimension.
-This results in a vector double the dimension of the input for each time step.}.
-After max-pooling the BiLSTM output over the sequence dimension we have a single 4096-dim vector corresponding to the tweet representation.
-This representation is finally fed to a single-layer fully-connected network with input size 4096, 512 hidden units, output size 6, and ReLU nonlinearities after the input and hidden layers.
-The output of the dense layer is a 6-dim logit vector for each example.
+\textbf{Dimensionalities}: By default the ELMo layer outputs a 1024-dim vector,
+which we then feed to a BiLSTM with output size 2048, resulting in a 4096-dim
+vector when concatenating forward and backward directions for each word
+of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read
+the input in opposite directions and whose outputs are concatenated at the
+hidden dimension. This results in a vector double the dimension of the input
+for each time step.}. After max-pooling the BiLSTM output over the sequence
+dimension, we obtain a single 4096-dim vector corresponding to the tweet
+representation. This representation is finally fed to a single-layer
+fully-connected network with input size 4096, 512 hidden units, output size 6,
+and a ReLU nonlinearity after the hidden layer. The output of the
+dense layer is a 6-dim logit vector for each input example.
 
-\textbf{Loss Function}: As this corresponds to a multiclass classification problem (predicting a single class for each example, with more than 2 classes to choose from), we used the Cross-Entropy Loss as implemented in PyTorch \cite{paszke2017automatic}.
+\textbf{Loss Function}: As this corresponds to a multiclass classification
+problem (predicting a single class for each example, with more than 2 classes to
+choose from), we used the Cross-Entropy Loss as implemented in PyTorch
+\cite{paszke2017automatic}.
 
-\textbf{Optimization}: We optimized the model with Adam \cite{DBLP:journals/corr/KingmaB14}, using a slanted triangular learning rate schedule \cite{howard2018universal}, with default hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate $\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained by multiplying the number of epochs ($10$) times the total number of batches, which for the training dataset corresponds to $2396$ batches of $64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.
+\textbf{Optimization}: We optimized the model with Adam
+\cite{DBLP:journals/corr/KingmaB14}, using default hyperparameters
+($\beta_1=0.9$, $\beta_2=0.999$, $\epsilon=10^{-8}$), following a slanted
+triangular learning rate schedule \cite{howard2018universal}, also with default
+hyperparameters ($cut\_frac=0.1$, $ratio=32$), and a maximum learning rate
+$\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained
+    by multiplying the number of epochs ($10$) times the total number of
+    batches, which for the training dataset corresponds to $2396$ batches of
+$64$ elements, and $1$ batch of $39$ elements, hence $2397\times10=23,970$.}.
 
-\textbf{Regularization}: we used a dropout layer \cite{srivastava2014dropout}, with probability of 0.5 after both the ELMo and the hidden fully-connected layer, and another one with probability of 0.1 after the max-pooling aggregation layer.
-We also reshuffled the training examples between epochs, resulting in a different batch for each iteration.
+\textbf{Regularization}: we used a dropout layer \cite{srivastava2014dropout},
+with probability of 0.5 after both the ELMo and the hidden fully-connected
+layer, and another one with probability of 0.1 after the max-pooling aggregation
+layer. We also reshuffled the training examples between epochs, resulting in a
+different batch for each iteration.
 
-\textbf{Model Selection}: To choose the best hyperparameter configuration we measured the classification accuracy on the validation set (trial set).
+\textbf{Model Selection}: To choose the best hyperparameter configuration we
+measured the classification accuracy on the validation (trial) set.
 
@@
 
-Once we found the best-performing configuration we trained 10 models using different random seeds, and tried averaging the output probabilities of all their possible $\sum_{k=1}^{10}{\binom{10}{k}}=1023$ combinations.
-As Figure \ref{fig:best_ensembles} shows, we empirically found that a specific combination of $6$ models yielded the best results ($70.52\%$), providing evidence for the fact that using a number of independent classifiers equal to the number of class labels provides the best results when doing average ensembling \cite{bonab2016theoretical}.
+Once we found the best-performing configuration we trained 10 models using
+different random seeds, and tried averaging the output probabilities of all
+their possible $\sum_{k=1}^{10}{\binom{10}{k}}=1023$ combinations. As Figure
+\ref{fig:best_ensembles} shows, we empirically found that a specific combination
+of $6$ models yielded the best results ($70.52\%$), providing evidence for the
+fact that using a number of independent classifiers equal to the number of class
+labels provides the best results when doing average ensembling
+\cite{bonab2016theoretical}.
 
@@
 
-We performed several experiments to gain insights on how the proposed model's performance interacts with the shared task's data.
-We performed an ablation study to see how some of the main hyperparameters affect performance, and an analysis of tweets containing hashtags and emoji to understand how these two types of tokens help the model predict the trigger-word's emotion.
-We also observed the effects of varying the amount of data used for training the model to evaluate whether it would be worthwhile to gather more training data.
+We performed several experiments to gain insights on how the proposed model's
+performance interacts with the shared task's data. We performed an ablation
+study to see how some of the main hyperparameters affect performance, and an
+analysis of tweets containing hashtags and emoji to understand how these two
+types of tokens help the model predict the trigger-word's emotion. We also
+observed the effects of varying the amount of data used for training the model
+to evaluate whether it would be worthwhile to gather more training data.
 
@@
 
-We performed an ablation study on a single model having obtained 69.23\% accuracy on the validation set. Results are summarized in Table \ref{table:ablation}.
+We performed an ablation study on a single model having obtained 69.23\%
+accuracy on the validation set. Results are summarized in Table
+\ref{table:ablation}.
 
@@
         obtained sentence representations by using the pooling method described
-        by \citet{howard2018universal}. ``LSTM hidden'' corresponds to the the
+        by \citet{howard2018universal}. ``LSTM hidden'' corresponds to the
         hidden dimension of the BiLSTM, ``POS emb dim'' to the dimension of the
@@
 
-We can observe that the architectural choice that had the greatest impact on our model was the ELMo layer, providing a $3.71\%$ boost in performance as compared to using GloVe pre-trained word embeddings.
+We can observe that the architectural choice that had the greatest impact on our
+model was the ELMo layer, providing a $3.71\%$ boost in performance as compared
+to using GloVe pre-trained word embeddings.
 
-We can further see that emoji also contributed significantly to the model's performance. In Section \ref{sub:effect_of_emoji_and_hashtags} we give some pointers to understanding why this is so.
+We can further see that emoji also contributed significantly to the model's
+performance. In Section \ref{sub:effect_of_emoji_and_hashtags} we give some
+pointers to understanding why this is so.
 
-Additionally, we tried using the concatenation of the max-pooled, average-pooled and last hidden states of the BiLSTM as the sentence representation, following \citet{howard2018universal}, but found out that this impacted performance negatively.
-We hypothesize this is due to tweets being too short for needing such a rich representation.
-Also, the size of the concatenated vector was $4096\times3=12,288$, which probably could not be properly exploited by the $512$-dimensional fully-connected layer.
+Additionally, we tried using the concatenation of the max-pooled, average-pooled
+and last hidden states of the BiLSTM as the sentence representation, following
+\citet{howard2018universal}, but found out that this impacted performance
+negatively. We hypothesize this is due to tweets being too short for needing
+such a rich representation. Also, the size of the concatenated vector was
+$4096\times3=12,288$, which probably could not be properly exploited by the
+$512$-dimensional fully-connected layer.
 
-Using a greater BiLSTM hidden size did not help the model, probably because of the reason mentioned early; the fully-connected layer was not big or deep enough to exploit the additional information.
-Similarly, using a smaller hidden size neither helped.
+Using a greater BiLSTM hidden size did not help the model, probably because of
+the reason mentioned early; the fully-connected layer was not big or deep enough
+to exploit the additional information. Similarly, using a smaller hidden size
+neither helped.
 
-We found that using 50-dim part-of-speech embeddings slightly improved results, which implies that better fine-tuning this hyperparameter, or using a better POS tagger could yield an even better performance.
+We found that using 50-dim part-of-speech embeddings slightly improved results,
+which implies that better fine-tuning this hyperparameter, or using a better POS
+tagger could yield an even better performance.
 
-Regarding optimization strategies, we also tried using SGD with different learning rates and a step-wise learning rate schedule as described by \citet{conneau2018}, but we found that doing this did not improve performance.
+Regarding optimization strategies, we also tried using SGD with different
+learning rates and a step-wise learning rate schedule as described by
+\citet{conneau2018}, but we found that doing this did not improve performance.
 
@@
 for the model to classify than those without. Hashtags also have a positive
-effect on classification performance, however it is less significant. This implies that emoji, and hashtags in a smaller degree, provide
-tweets with a context richer in sentiment information, allowing the model to
-better guess the emotion of the \textit{trigger-word}.
+effect on classification performance, however it is less significant. This
+implies that emoji, and hashtags in a smaller degree, provide tweets with a
+context richer in sentiment information, allowing the model to better guess the
+emotion of the \textit{trigger-word}.
 
-Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on classification performance.
+Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on
+classification performance.
 
@@
 
-It is clear some emoji strongly contribute to improving prediction quality.
-The most interesting ones are \texttt{mask}, \texttt{rage}, and \texttt{cry}, which significantly increase accuracy.
-Further, contrary to intuition, the \texttt{sob} emoji contributes less than \texttt{cry}, despite representing a stronger emotion.
-This is probably due to \texttt{sob} being used for depicting a wider spectrum of emotions.
+It is clear some emoji strongly contribute to improving prediction quality. The
+most interesting ones are \texttt{mask}, \texttt{rage}, and \texttt{cry}, which
+significantly increase accuracy. Further, contrary to intuition, the
+\texttt{sob} emoji contributes less than \texttt{cry}, despite representing a
+stronger emotion. This is probably due to \texttt{sob} being used for depicting
+a wider spectrum of emotions.
 
-Finally, not all emoji are beneficial for this task.
-When removing \texttt{sweat\_smile} and \texttt{confused} accuracy increased, probably because they represent emotions other than the ones being predicted.
-%From observing Table \ref{table:emoji_label} it seems likely that tweets containing \texttt{mask} would be highly predictive of the \texttt{disgust} class, since more than $92\%$ of the tweets containing such emoji belong to that class, however it less clear why \texttt{rage}, and \texttt{cry} contribute similarly, despite being more evenly distributed among classes.
+Finally, not all emoji are beneficial for this task. When removing
+\texttt{sweat\_smile} and \texttt{confused} accuracy increased, probably because
+they represent emotions other than the ones being predicted.
 
-%This means that they are highly representative of the the class the tweet belongs to, and would be good choices for discriminating new tweets in case we wanted to expand the dataset.
+%From observing Table \ref{table:emoji_label} it seems likely that tweets
+%containing \texttt{mask} would be highly predictive of the \texttt{disgust}
+%class, since more than $92\%$ of the tweets containing such emoji belong to
+%that class, however it less clear why \texttt{rage}, and \texttt{cry}
+%contribute similarly, despite being more evenly distributed among classes.
 
-%For example, by observing Table \ref{table:emoji_label} we can infer that  \texttt{mask}, \texttt{rage}, and \texttt{cry}   \texttt{disgust}, \texttt{rage}, and \texttt{sad} classes respectively.
+%This means that they are highly representative of the the class the tweet
+%belongs to, and would be good choices for discriminating new tweets in case we
+%wanted to expand the dataset.
+
+%For example, by observing Table \ref{table:emoji_label} we can infer that
+%\texttt{mask}, \texttt{rage}, and \texttt{cry}   \texttt{disgust},
+%\texttt{rage}, and \texttt{sad} classes respectively.
 
@@
 
-As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data in which our model was trained consistently increased validation accuracy and validation macro F1 score.
-The trend shows that having more training data would likely further increase the model's performance.
+As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data in
+which our model was trained consistently increased validation accuracy and
+validation macro F1 score. The trend shows that having more training data would
+likely further increase the model's performance.
 
@@
 
-We described the model that got second place in the WASSA 2018 Implicit Emotion Shared Task.
-Despite its simplicity, and low amount of dependencies on libraries and external features, it performed almost as well as the system that obtained the first place.
+We described the model that got second place in the WASSA 2018 Implicit Emotion
+Shared Task. Despite its simplicity, and low amount of dependencies on libraries
+and external features, it performed almost as well as the system that obtained
+the first place.
 
-Our ablation study revealed that our hyperparameters were indeed quite well-tuned for the task, which agrees with the good results obtained in the official submission.
-However, the ablation study also showed that increased performance can be obtained by incorporating POS embeddings as additional inputs.
-Further experiments are required to accurately measure the impact that this additional input may have on the results.
-We also think the performance can be boosted by making the architecture more complex, concretely, by using a BiLSTM with multiple layers and skip connections in a way akin to \cite{peters2018deep}, or by making the fully-connected network bigger and deeper.
+Our ablation study revealed that our hyperparameters were indeed quite
+well-tuned for the task, which agrees with the good results obtained in the
+official submission. However, the ablation study also showed that increased
+performance can be obtained by incorporating POS embeddings as additional
+inputs. Further experiments are required to accurately measure the impact that
+this additional input may have on the results. We also think the performance can
+be boosted by making the architecture more complex, concretely, by using a
+BiLSTM with multiple layers and skip connections in a way akin to
+\cite{peters2018deep}, or by making the fully-connected network bigger and
+deeper.
 
-Finally, our studies regarding the importance of hashtags and emoji in the classification showed that both of them seem to contribute significantly to the performance, although in different measures.
+Finally, our studies regarding the importance of hashtags and emoji in the
+classification showed that both of them seem to contribute significantly to the
+performance, although in different measures.
+
+\section{Acknowledgements}%
+\label{sec:acknowledgements}
+
+We thank the anonymous reviewers for their reviews and suggestions. The first
+author is partially supported by the Japanese Government MEXT Scholarship.
 

### 2018-09-01 paper/paper.tex
[paper] Improve tables, begin writing Error Analysis subsection

@@
 \usepackage{blindtext}
+\usepackage{booktabs}
 \usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
@@
     \begin{tabular}{ll}
-
+        \toprule
         \textbf{Original} & \textbf{Replacement} \\
-        \hline
-        \hline
+        \midrule
         \texttt{\footnotesize[\#TRIGGERWORD\#]} & \texttt{\footnotesize\_\_TRIGGERWORD\_\_} \\
@@
         \texttt{\footnotesize http://url.removed} & \texttt{\footnotesize\_\_URL\_\_} \\
+        \bottomrule
 
@@
 
-\textbf{Dimensionalities}: By default the ELMo layer outputs a 1024-dim vector,
-which we then feed to a BiLSTM with output size 2048, resulting in a 4096-dim
-vector when concatenating forward and backward directions for each word
-of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read
-the input in opposite directions and whose outputs are concatenated at the
-hidden dimension. This results in a vector double the dimension of the input
+\textbf{Dimensionalities}: By default the ELMo layer outputs a 1024-dimensional
+vector, which we then feed to a BiLSTM with output size 2048, resulting in a
+4096-dimensional vector when concatenating forward and backward directions for
+each word of the sequence\footnote{A BiLSTM is composed of two separate LSTMs
+that read the input in opposite directions and whose outputs are concatenated at
+the hidden dimension. This results in a vector double the dimension of the input
 for each time step.}. After max-pooling the BiLSTM output over the sequence
-dimension, we obtain a single 4096-dim vector corresponding to the tweet
+dimension, we obtain a single 4096-dimensional vector corresponding to the tweet
 representation. This representation is finally fed to a single-layer
 fully-connected network with input size 4096, 512 hidden units, output size 6,
-and a ReLU nonlinearity after the hidden layer. The output of the
-dense layer is a 6-dim logit vector for each input example.
+and a ReLU nonlinearity after the hidden layer. The output of the dense layer is
+a 6-dimensional logit vector for each input example.
 
@@
 
+        \toprule
         \textbf{Variation} & \textbf{Accuracy (\%)} & $\bf{\Delta}$\textbf{\%} \\
-        \hline
-        \hline
+        \midrule[0.08em]
         Submitted          & \textbf{69.23}         & -                        \\
-        \hline
+        \midrule[0.05em]
         No emoji           & 68.36                  & - 0.87                   \\
-        \hline
+        \midrule[0.05em]
         No ELMo            & 65.52                  & - 3.71                   \\
-        \hline
+        \midrule[0.05em]
         Concat Pooling     & 68.47                  & - 0.76                   \\
-        \hline
+        \midrule[0.05em]
         LSTM hidden=4096   & 69.10                  & - 0.13                   \\
@@
         LSTM hidden=512    & 68.43                  & - 0.80                   \\
-        \hline
+        \midrule[0.05em]
         POS emb dim=100    & 68.99                  & - 0.24                   \\
@@
         POS emb dim=25     & 69.21                  & - 0.02                   \\
-        \hline
+        \midrule[0.05em]
         SGD optim lr=1     & 64.33                  & - 4.90                   \\
@@
         SGD optim lr=0.001 & 30.49                  & - 38.74                  \\
+        \bottomrule
 
@@
 Using a greater BiLSTM hidden size did not help the model, probably because of
-the reason mentioned early; the fully-connected layer was not big or deep enough
+the reason mentioned earlier; the fully-connected layer was not big or deep enough
 to exploit the additional information. Similarly, using a smaller hidden size
@@
 
-We found that using 50-dim part-of-speech embeddings slightly improved results,
+We found that using 50-dimensional part-of-speech embeddings slightly improved results,
 which implies that better fine-tuning this hyperparameter, or using a better POS
@@
 
-    \caption{Dropout Ablation. }
+    \caption{Dropout Ablation.}
     \vspace{-0.4cm}
@@
 
+\subsection{Error Analysis}
+
+Figure~\ref{fig:confusion_matrix} shows the confusion matrix of the model
+evaluated on the test set, and Table~\ref{table:classification_report} the
+corresponding classification report.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=\columnwidth]{images/confusion_matrix.pdf}
+
+    \caption{Confusion Matrix (Test Set).}
+    % \vspace{-0.4cm}
+
+    % \caption*{
+    %     \footnotesize Rows correspond to the dropout applied both after
+    %     the ELMo layer (word encoding layer) and after the fully-connected network's
+    %     hidden layer, while columns correspond to the dropout applied after the
+    %     max-pooling operation (sentence encoding layer.)}
+
+    \label{fig:confusion_matrix}
+\end{figure}
+
+\begin{table}[htpb]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{cccc}
+
+        \toprule
+                          & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
+        \midrule
+        \texttt{anger}    & 0.643     & 0.601  & 0.621    \\
+        \texttt{disgust}  & 0.703     & 0.661  & 0.682    \\
+        \texttt{fear}     & 0.742     & 0.721  & 0.732    \\
+        \texttt{joy}      & 0.762     & 0.805  & 0.783    \\
+        \texttt{sad}      & 0.685     & 0.661  & 0.673    \\
+        \texttt{surprise} & 0.627     & 0.705  & 0.663    \\
+        \midrule
+        Average           & 0.695     & 0.695  & 0.694    \\
+        \bottomrule
+    \end{tabular}
+
+    \caption{Classification Report (Test Set).}
+
+\label{table:classification_report}
+\end{table}
+
 \subsection{Effect of Emoji and Hashtags}%
@@
 
+        \toprule
         \textbf{} & \textbf{Present} & \textbf{Not Present} \\
-        \hline
-        \hline
+        \midrule
         Emoji     & 4805 (76.6\%)    & 23952 (68.0\%)       \\
         Hashtags  & 2122 (70.5\%)    & 26635 (69.4\%)       \\
+        \bottomrule
 
@@
 
-Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on
-classification performance.
-
 \begin{table}[!h]
@@
     \scriptsize
-        \begin{tabular}{lr|rc|rc|r}
+        \begin{tabular}{lrrcrcr}
 
-           \multirow{2}{*}{\textbf{Emoji alias}} & \multirow{2}{*}{\textbf{N}} & \multicolumn{2}{c|}{\textbf{emoji}} & \multicolumn{2}{c|}{\textbf{no-emoji}} &
+            \toprule
+           \multirow{2}{*}{\textbf{Emoji alias}} & \multirow{2}{*}{\textbf{N}\hspace{0.1cm}} & \multicolumn{2}{c}{\textbf{emoji}} & \multicolumn{2}{c}{\textbf{no-emoji}} &
             \multirow{2}{*}{$\bf{\Delta}$\textbf{\%}} \\
-            \cline{3-6}
+            \cmidrule(lr){3-4}
+            \cmidrule(lr){5-6}
                                 &     & \#   & \% & \# & \% & \\
-          \hline
-          \hline
-          \texttt{mask}         & 163 & 154 & 94.48 & 134 & 82.21 & - 12.27 \\
-          \texttt{two\_hearts}  & 87  & 81  & 93.10 & 77  & 88.51 & - 4.59  \\
-          \texttt{heart\_eyes}  & 122 & 109 & 89.34 & 103 & 84.43 & - 4.91  \\
-          \texttt{heart}        & 267 & 237 & 88.76 & 235 & 88.01 & - 0.75  \\
-          \texttt{rage}         & 92  & 78  & 84.78 & 66  & 71.74 & - 13.04 \\
-          \texttt{cry}          & 116 & 97  & 83.62 & 83  & 71.55 & - 12.07 \\
-          \texttt{sob}          & 490 & 363 & 74.08 & 345 & 70.41 & - 3.67  \\
-          \texttt{unamused}     & 167 & 121 & 72.46 & 116 & 69.46 & - 3.00  \\
-          \texttt{weary}        & 204 & 140 & 68.63 & 139 & 68.14 & - 0.49  \\
-          \texttt{joy}          & 978 & 649 & 66.36 & 629 & 64.31 & - 2.05  \\
-          \texttt{sweat\_smile} & 111 & 73  & 65.77 & 75  & 67.57 & 1.80 \\
-          \texttt{confused}     & 77  & 46  & 59.74 & 48  & 62.34 & 2.60 \\
+            \midrule[0.08em]
+            \texttt{mask}       & 163   \hspace{0.1cm} & 154 & 94.48 \hspace{0.1cm} & 134 & 82.21 & - 12.27 \\
+          \texttt{two\_hearts}  & 87    \hspace{0.1cm} & 81  & 93.10 \hspace{0.1cm} & 77  & 88.51 & - 4.59  \\
+          \texttt{heart\_eyes}  & 122   \hspace{0.1cm} & 109 & 89.34 \hspace{0.1cm} & 103 & 84.43 & - 4.91  \\
+          \texttt{heart}        & 267   \hspace{0.1cm} & 237 & 88.76 \hspace{0.1cm} & 235 & 88.01 & - 0.75  \\
+          \midrule[0.001em]
+          \texttt{rage}         & 92    \hspace{0.1cm} & 78  & 84.78 \hspace{0.1cm} & 66  & 71.74 & - 13.04 \\
+          \texttt{cry}          & 116   \hspace{0.1cm} & 97  & 83.62 \hspace{0.1cm} & 83  & 71.55 & - 12.07 \\
+          \texttt{sob}          & 490   \hspace{0.1cm} & 363 & 74.08 \hspace{0.1cm} & 345 & 70.41 & - 3.67  \\
+          \texttt{unamused}     & 167   \hspace{0.1cm} & 121 & 72.46 \hspace{0.1cm} & 116 & 69.46 & - 3.00  \\
+          \midrule[0.001em]
+          \texttt{weary}        & 204   \hspace{0.1cm} & 140 & 68.63 \hspace{0.1cm} & 139 & 68.14 & - 0.49  \\
+          \texttt{joy}          & 978   \hspace{0.1cm} & 649 & 66.36 \hspace{0.1cm} & 629 & 64.31 & - 2.05  \\
+          \texttt{sweat\_smile} & 111   \hspace{0.1cm} & 73  & 65.77 \hspace{0.1cm} & 75  & 67.57 & 1.80 \\
+          \texttt{confused}     & 77    \hspace{0.1cm} & 46  & 59.74 \hspace{0.1cm} & 48  & 62.34 & 2.60 \\
+          \bottomrule
 
@@
 
-It is clear some emoji strongly contribute to improving prediction quality. The
+Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on
+classification performance. It is clear some emoji strongly contribute to improving prediction quality. The
 most interesting ones are \texttt{mask}, \texttt{rage}, and \texttt{cry}, which
@@
     \includegraphics[width=\columnwidth]{images/acc_vs_tdp_variation.eps}
-    \caption{Validation accuracy and macro F1 score vs. proportion of the training data.}
+    \caption{Effect of the amount of training data on classification performance.}
     \label{fig:data_amt_vs_acc}

### 2018-09-01 paper/paper.tex
[paper] Add brief error analysis, expand conclusion, add pca figure

@@
 Network BiLSTM for enriching word representations with context, a max-pooling
-operation for creating sentence representations from them, and a
-Dense Layer for projecting the sentence representations into label space. Our
-official submission was obtained by ensembling 6 of these models initialized
-with different random seeds.
+operation for creating sentence representations from them, and a Dense Layer for
+projecting the sentence representations into label space. Our official
+submission was obtained by ensembling 6 of these models initialized with
+different random seeds. The code for replicating this paper is available at
+\url{https://github.com/jabalazs/implicit_emotion}.
 
@@
 $\eta_{max}=0.001$, over $T=23,970$ iterations\footnote{This number is obtained
-    by multiplying the number of epochs ($10$) times the total number of
+    by multiplying the number of epochs ($10$), times the total number of
     batches, which for the training dataset corresponds to $2396$ batches of
@@
 Once we found the best-performing configuration we trained 10 models using
-different random seeds, and tried averaging the output probabilities of all
-their possible $\sum_{k=1}^{10}{\binom{10}{k}}=1023$ combinations. As Figure
+different random seeds, and tried averaging the output class probabilities of
+all their possible $\sum_{k=1}^{9}{\binom{9}{k}}=511$ combinations. As Figure
 \ref{fig:best_ensembles} shows, we empirically found that a specific combination
@@
 
-Figure~\ref{fig:confusion_matrix} shows the confusion matrix of the model
+Figure~\ref{fig:confusion_matrix} shows the confusion matrix of a single model
 evaluated on the test set, and Table~\ref{table:classification_report} the
-corresponding classification report.
+corresponding classification report. In general, we confirm what
+\newcite{klinger2018iest} report: \texttt{anger} was the most difficult class to
+predict, followed by \texttt{surprise}, whereas \texttt{joy}, \texttt{fear}, and
+\texttt{disgust} are the better performing ones.
+
+To observe whether any particular pattern arose from the sentence
+representations encoded by our model, we projected them into 3d space through
+Principal Component Analysis (PCA), and were surprised to find that 2 clearly
+defined clusters emerged (see Figure \ref{fig:pca}), one containing the majority
+of datapoints, and another containing \texttt{joy} tweets exclusively. Upon
+further exploration we also found that the smaller cluster was composed only by
+tweets containing the pattern \texttt{un~\_\_TRIGGERWORD\_\_}, and further,
+that all of them were correctly classified.
+
+It is also worth mentioning that there are 5827 tweets in the training set with
+this pattern. Of these, 5822 (99.9\%) correspond to the label \texttt{joy}. We
+observe a similar trend on the test set; 1115 of the 1116 tweets having the
+\texttt{un~\_\_TRIGGERWORD\_\_} pattern correspond to \texttt{joy} tweets. We
+hypothesize this is the reason why the model learned this pattern as a strong
+discriminating feature.
+
+Finally, the only tweet in the test set that contained this pattern and did not
+belong to the \texttt{joy} class, originally had \textit{unsurprised} as its
+triggerword\footnote{We manually searched for the original tweet.}, and
+unsurprisingly, was misclassified.
 
@@
 \end{table}
+
+\subsection{Effect of the Amount of Training Data}
+
+As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data with
+which our model was trained consistently increased validation accuracy and
+validation macro F1 score. The trend suggests that the proposed model is
+expressive enough to learn from more data, and is not overfitting the training
+set.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=\columnwidth]{images/acc_vs_tdp_variation.eps}
+    \caption{Effect of the amount of training data on classification performance.}
+    \label{fig:data_amt_vs_acc}
+\end{figure}
 
@@
 
-\subsection{Effect of the Amount of Training Data}
-
-As Figure \ref{fig:data_amt_vs_acc} shows, increasing the amount of data in
-which our model was trained consistently increased validation accuracy and
-validation macro F1 score. The trend shows that having more training data would
-likely further increase the model's performance.
-
 \begin{figure}[!h]
     \centering
-    \includegraphics[width=\columnwidth]{images/acc_vs_tdp_variation.eps}
-    \caption{Effect of the amount of training data on classification performance.}
-    \label{fig:data_amt_vs_acc}
+    \includegraphics[width=\columnwidth]{images/pca.pdf}
+
+    \caption{3d Projection of the Test Sentence Representations.}
+    \label{fig:pca}
 \end{figure}
@@
 
+We also showed that, what was probably an annotation artifact, the
+\texttt{un~\_\_TRIGGERWORD\_\_} pattern, resulted in increased performance for
+the \texttt{joy} label. This pattern was probably originated by a heuristic
+naïvely replacing the ocurrence of \textit{happy} by the trigger-word indicator.
+We think the dataset could be improved by replacing the word \textit{unhappy},
+in the original examples, by \texttt{\_\_TRIGGERWORD\_\_} instead of
+\texttt{un~\_\_TRIGGERWORD\_\_}, and labeling it as \texttt{sad}, or
+\texttt{angry}, instead of \texttt{joy}.
+
+% To prevent future models from mistakenly learning this
+% feature as a strong indicator for such label, it would be wise to add more
+% examples from other classes cotaining the same pattern, or removing them
+% entirely and rebalancing the dataset.
+
 Finally, our studies regarding the importance of hashtags and emoji in the

### 2018-09-03 run.py
Fix logging message displaying incorrect information in test mode

@@ -138,8 +138,9 @@ def main():
     if hp.write_mode != 'NONE':
         logger.write_hyperparams()
 
-    print(f'Running experiment {logger.model_hash}. Hyperparameters and '
-          f'checkpoints will be saved in {logger.run_savepath}')
+    if not hp.test:
+        print(f'Running experiment {logger.model_hash}. Hyperparameters and '
+              f'checkpoints will be saved in {logger.run_savepath}')
 
     torch.manual_seed(hp.seed)
     torch.cuda.manual_seed_all(hp.seed)  # silently ignored if there are no GPUs

### 2018-09-03 scripts/preprocess.sh
Add code for creating dataset without trigger-words
This was suggested by one of the reviewers. We found that removing
triggerwords cause a drop of 0.66% absolute performance (validation
accuracy) on our best model.

@@ -82,11 +82,15 @@ echo "Created $CLEAN_TRAIN_DATA_PATH and $TRAIN_LABELS_PATH"
 
 echo
 
+echo "Creating Train dataset with no Triggerwords"
+cat $CLEAN_TRAIN_DATA_PATH | awk '{gsub("__TRIGGERWORD__ ", "", $0); print $0}' > $CLEAN_TRAIN_NO_TRIGGERWORD_DATA_PATH
+echo
+
 echo "Creating Train dataset with no emoji"
 cat $CLEAN_TRAIN_DATA_PATH | ./utils/remove_emojis.py > $CLEAN_TRAIN_NO_EMOJIS_DATA_PATH
 echo "Created $CLEAN_TRAIN_NO_EMOJIS_DATA_PATH"
 
-echo
+echo "################################"
 
 ################################################################################
 #                                     DEV                                      #
@@ -112,11 +116,15 @@ echo "Created $CLEAN_DEV_DATA_PATH and $DEV_LABELS_PATH"
 
 echo
 
+echo "Creating Dev dataset with no Triggerwords"
+cat $CLEAN_DEV_DATA_PATH | awk '{gsub("__TRIGGERWORD__ ", "", $0); print $0}' > $CLEAN_DEV_NO_TRIGGERWORD_DATA_PATH
+echo
+
 echo "Creating Dev dataset with no emoji"
 cat $CLEAN_DEV_DATA_PATH | ./utils/remove_emojis.py > $CLEAN_DEV_NO_EMOJIS_DATA_PATH
 echo "Created $CLEAN_DEV_NO_EMOJIS_DATA_PATH"
 
-echo
+echo "################################"
 ################################################################################
 #                                     TEST                                     #
 ################################################################################
@@ -128,6 +136,11 @@ preprocess $TEST_DATA_PATH $CLEAN_TEST_DATA_PATH
 
 cat $TEST_DATA_PATH | awk '{if ( $0 ~ /#TRIGGERWORD#/ ) {print $1}}' > $TEST_LABELS_PATH
 echo "Created $CLEAN_TEST_DATA_PATH and $TEST_LABELS_PATH"
+
+echo
+
+echo "Creating Test dataset with no Triggerwords"
+cat $CLEAN_TEST_DATA_PATH | awk '{gsub("__TRIGGERWORD__ ", "", $0); print $0}' > $CLEAN_TEST_NO_TRIGGERWORD_DATA_PATH
 echo
 
 echo "Creating Test dataset with no emoji"

### 2018-09-03 src/config.py

@@ -40,6 +40,10 @@ TRAIN_POS = os.path.join(PREPROCESSED_DATA_PATH, 'train.csv.tagged.pos')
 DEV_POS = os.path.join(PREPROCESSED_DATA_PATH, 'dev.csv.tagged.pos')
 TEST_POS = os.path.join(PREPROCESSED_DATA_PATH, 'test.csv.tagged.pos')
 
+TRAIN_NO_TRIGGERWORD = os.path.join(PREPROCESSED_DATA_PATH, 'train_no_triggerword.csv')
+DEV_NO_TRIGGERWORD = os.path.join(PREPROCESSED_DATA_PATH, 'dev_no_triggerword.csv')
+TEST_NO_TRIGGERWORD = os.path.join(PREPROCESSED_DATA_PATH, 'test_no_triggerword.csv')
+
 
 # MAPPINGS
 embedding_dict = {
@@ -62,6 +66,11 @@ corpora_dict = {
         'dev': DEV_EMOJI,
         'test': TEST_EMOJI
     },
+    'iest_no_triggerword': {
+        'train': TRAIN_NO_TRIGGERWORD,
+        'dev': DEV_NO_TRIGGERWORD,
+        'test': TEST_NO_TRIGGERWORD
+    }
 }
 
 pos_corpora_dict = {

### 2018-09-03 run.py
Update readme with info on testing, improve test result filenames

@@ -287,13 +287,13 @@ def main():
 
         probs = np_softmax(eval_dict['output'])
         probs_filepath = os.path.join(ext_experiment_path,
-                                      'test_probs.csv')
+                                      'test_probabilities.csv')
         np.savetxt(probs_filepath, probs,
                    delimiter=',', fmt='%.8f')
         print(f'Saved prediction probs in {probs_filepath}')
 
         labels_filepath = os.path.join(ext_experiment_path,
-                                       'predictions.txt')
+                                       'test_predictions.txt')
         labels = [label + '\n' for label in eval_dict['labels']]
         with open(labels_filepath, 'w', encoding='utf-8') as f:
             f.writelines(labels)
